---
title: "Week 2 Peer Review"
output: html_document
---

```{r setup, include=FALSE}

######DO NOT MODIFY. This will load required packages and data.

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
cces <- drop_na(read_csv(url("https://www.dropbox.com/s/ahmt12y39unicd2/cces_sample_coursera.csv?raw=1")))
```

Your objective is to replicate these figures, created using the Cooperative Congressional Election Study data. These figures are similar to those we completed in the lecture videos.

### Put your name here: Minh Q Pham



## Exercise 1

Hints:

* For the x-axis, use the variable "ideo5".

* Make sure you recode the data for the "ideo5" variable to generate the correct names for the x-axis. You will want to consult the codebook.

* Use the fill aesthetic to have R fill in the bars. You do not need to set the colors manually.

* Use guides() to drop the legend.

* Make sure the axis labels and figure title are correct.


```{r exercise1,echo=FALSE}
cces$ideo5 <- recode(cces$ideo5, `1`="Very Liberal", `2`="Liberal", `3`="Moderate", `4`="Conservative", `5`= "Very Conservative" )
ggplot(cces, aes(x=ideo5, fill=ideo5)) + geom_bar() + 
  labs(x="Ideology", y="Count", title = "Respondent Ideology") +guides(fill=FALSE)
```

## Exercise 2

Hints:

* For the x-axis, use the variable "pew_religimp".

* Make sure you recode the data for the "pew_religimp" variable to generate the correct labels for the x-axis. You will want to consult the codebook.

* Rename the column for Ideology to make sure the first letter is upper-case (to make the legend appear correctly).

* Use the fill aesthetic to have R fill in the bars. You do not need to set the colors manually.

* Make sure the axis labels and figure title are correct.

```{r exercise2,echo=FALSE}
cces$pew_religimp <- recode(cces$pew_religimp, `1`="Very imp.", `2`= "Somewhat imp.", `3`= "Not too imp.", `4`="Not at all imp.")
cces <- cces %>% rename(Ideology=ideo5)
ggplot(cces, aes(x=pew_religimp, fill= Ideology)) + geom_bar() + 
  labs(x="Importance of Religion", y="Count", title = "Importance of Religion by Ideology")
```

## Exercise 3

Instructions:

* For this visualization, you are creating your own data for practice.

* Create a tibble/data frame with three columns: Semester, Student, and Grade.

* There should be six semesters and three students (Amanda, Betty, and Carol)

* Create grades for the students using the runif() command, with values between 80 and 100. Hint: you'll need 18 grades total.

The figure should look approximately like this (your vaules will be slightly different):


```{r exercise3,echo=FALSE}
Semester <- rep(c(1,2,3,4,5,6),3)
Amanda <- rep("Amanda",6)
Betty <- rep("Betty", 6)
Carol <- rep("Carol", 6)
Student <- c(Amanda, Betty, Carol)

Grade <- runif(18, 80, 100)

ex3data <- tibble(Semester,Student,Grade)

ggplot(ex3data, aes(x=Semester, y= Grade, color=Student)) + geom_line() +
  labs(title = "Student Grades by Semester") + facet_wrap(~Student) +
  guides(color=FALSE)
  
```
