---
title: "Volatility Overload"
author: "Minh Pham & Chandler Clemons"
date: \today
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
header-includes:
- \usepackage{longtable,booktabs,multirow}
- \usepackage[justification=centering]{caption}
- \usepackage{float}
- \floatplacement{figure}{H}
linestretch: 2.0
longtable: yes
geometry: left=2.0cm,right=2.0cm,top=2cm,bottom=2cm
subtitle: How Market Volatility Affects Self-Reported Well-Being
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=55),tidy=TRUE, warning = F,message = F)
```


```{r,include=FALSE, echo=FALSE}
library(quantmod)
library(devtools)
library(readr)
library(pscl)
library(spdep)
library(MASS)
library(readxl)
library(lodown)
library(tidyverse)
library(survey)
library(zoo)
library(dplyr)
library(srvyr)
library(ggplot2)
library(sjstats)
library(quantmod)
library(multiwayvcov)
library(InformationValue)
library(lmtest)
library(mpath)
library(rsq)
library(stargazer)
library(gtsummary)
library(gt)
library(glue)
library(effects)
library(reshape2)
library(ROSE)
library(countreg)
library(mfx)
library(interactions)
library(Hmisc)
```


```{r, , include=FALSE, echo=FALSE}
# cpsasec_cat <-
#     get_catalog( "cpsasec" ,
#         output_dir = file.path( path.expand( "~" ) , "CPSASEC" ) )

cpsasec_df13 <- 
    readRDS( file.path( path.expand( "~" ) , "CPSASEC" , "2013 cps asec.rds" ) )
cpsasec_df14 <- 
    readRDS( file.path( path.expand( "~" ) , "CPSASEC" , "2014 cps asec.rds" ) )
cpsasec_df15 <- 
    readRDS( file.path( path.expand( "~" ) , "CPSASEC" , "2015 cps asec.rds" ) )
cpsasec_df16 <- 
    readRDS( file.path( path.expand( "~" ) , "CPSASEC" , "2016 cps asec.rds" ) )
cpsasec_df17 <- 
    readRDS( file.path( path.expand( "~" ) , "CPSASEC" , "2017 cps asec.rds" ) )


cpsasec_df13$iyear <- rep(2013,nrow(cpsasec_df13)) # make new column
cpsasec_df14$iyear <- rep(2014,nrow(cpsasec_df14)) # make new column
cpsasec_df15$iyear <- rep(2015,nrow(cpsasec_df15)) # make new column
cpsasec_df16$iyear <- rep(2016,nrow(cpsasec_df16)) # make new column
cpsasec_df17$iyear <- rep(2017,nrow(cpsasec_df17)) # make new column


cpsasec_df1314 <- intersect(colnames(cpsasec_df13), colnames(cpsasec_df14))

cpsasec_df1314 <- rbind(cpsasec_df13[ , cpsasec_df1314], cpsasec_df14[ , cpsasec_df1314])

cpsasec_df131415 <- intersect(colnames(cpsasec_df1314), colnames(cpsasec_df15))

cpsasec_df131415 <- rbind(cpsasec_df1314[ , cpsasec_df131415], cpsasec_df15[ , cpsasec_df131415])

cpsasec_df13141516 <- intersect(colnames(cpsasec_df131415), colnames(cpsasec_df16))

cpsasec_df13141516 <- rbind(cpsasec_df131415[ , cpsasec_df13141516], cpsasec_df16[ , cpsasec_df13141516])

cpsasec_df1314151617 <- intersect(colnames(cpsasec_df13141516), colnames(cpsasec_df17))

cpsasec_df1314151617 <- rbind(cpsasec_df13141516[ , cpsasec_df1314151617], cpsasec_df17[ , cpsasec_df1314151617])

#write_csv(cpsasec_df1314151617, '/Users/minhpham1609/Desktop/cpsasec_df1314151617.csv')
```


```{r, , include=FALSE, echo=FALSE}
cpsasec_df1314151617$state_name =
            factor(
                cpsasec_df1314151617$gestfips ,
                levels = 
                    c(1L, 2L, 4L, 5L, 6L, 8L, 9L, 10L, 
                    11L, 12L, 13L, 15L, 16L, 17L, 18L, 
                    19L, 20L, 21L, 22L, 23L, 24L, 25L, 
                    26L, 27L, 28L, 29L, 30L, 31L, 32L, 
                    33L, 34L, 35L, 36L, 37L, 38L, 39L, 
                    40L, 41L, 42L, 44L, 45L, 46L, 47L, 
                    48L, 49L, 50L, 51L, 53L, 54L, 55L, 
                    56L) ,
                labels =
                    c("Alabama", "Alaska", "Arizona", "Arkansas", "California", 
                    "Colorado", "Connecticut", "Delaware", "District of Columbia", 
                    "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", 
                    "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", 
                    "Massachusetts", "Michigan", "Minnesota", "Mississippi", "Missouri", 
                    "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", 
                    "New Mexico", "New York", "North Carolina", "North Dakota", "Ohio", 
                    "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", 
                    "South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", 
                    "Washington", "West Virginia", "Wisconsin", "Wyoming")
            ) 



cpsasec_dfs = cpsasec_df1314151617 %>% select(state_name,div_yn,iyear, a_sex, a_age, htotval, marsupwt, a_hga, prdtrace, hsup_wgt)  

#htotval : total household income
#PRDTRACE: race
#A_HGA: Educational attainment

cpsasec_dfs$ownstock = ifelse( cpsasec_dfs$div_yn ==1 , 1, ifelse(cpsasec_dfs$div_yn==2, 0, NA))

cpsasec_dfs$a_hga=ifelse( cpsasec_dfs$a_hga ==31 , 1, ifelse(cpsasec_dfs$a_hga==32, 1, ifelse(cpsasec_dfs$a_hga==33,1, ifelse(cpsasec_dfs$a_hga==34,1, ifelse(cpsasec_dfs$a_hga==35,1, ifelse(cpsasec_dfs$a_hga==36,1, ifelse(cpsasec_dfs$a_hga==37,1, ifelse(cpsasec_dfs$a_hga==38,1, ifelse(cpsasec_dfs$a_hga==39,2, ifelse(cpsasec_dfs$a_hga==40,3, ifelse(cpsasec_dfs$a_hga==41,3, ifelse(cpsasec_dfs$a_hga==42,3, ifelse(cpsasec_dfs$a_hga==43,4, ifelse(cpsasec_dfs$a_hga==44,4, ifelse(cpsasec_dfs$a_hga==45,4, ifelse(cpsasec_dfs$a_hga==46,4, NA))))))))))))))))

cpsasec_dfs$a_age= ifelse( cpsasec_dfs$a_age %in% 18:24 , 1 , ifelse( cpsasec_dfs$a_age %in% 25:34 , 2, ifelse( cpsasec_dfs$a_age %in% 35:44 , 3, ifelse( cpsasec_dfs$a_age %in% 45:54 , 4, ifelse( cpsasec_dfs$a_age %in% 55:64 , 5, ifelse( cpsasec_dfs$a_age %in% 65:85 , 6, NA ))))))

cpsasec_dfs$htotval= ifelse( cpsasec_dfs$htotval %in% 0:15000 , 1, ifelse(cpsasec_dfs$htotval %in% 15000:25000 , 2, ifelse(cpsasec_dfs$htotval %in% 25000:35000 , 3, ifelse(cpsasec_dfs$htotval %in% 35000:50000 , 4, ifelse(cpsasec_dfs$htotval %in% 50000:max(cpsasec_dfs$htotval) , 5,  NA)))))

cpsasec_dfs$prdtrace= ifelse( cpsasec_dfs$prdtrace ==1 , 1, ifelse( cpsasec_dfs$prdtrace ==2 , 2, ifelse( cpsasec_dfs$prdtrace ==3 , 3, ifelse( cpsasec_dfs$prdtrace ==4 , 4, ifelse( cpsasec_dfs$prdtrace ==5, 5, ifelse( cpsasec_dfs$prdtrace %in% 6:26, 6, NA))))))


cpsasec_dfs=na.omit(cpsasec_dfs)
colSums(is.na(cpsasec_dfs))

    
cpsasec_dfs$gender=as.factor(recode(cpsasec_dfs$a_sex,'1'='male','2'='female',.missing='NA',.default = 'NA'))

cpsasec_dfs$calculated_education <- as.factor(recode(cpsasec_dfs$a_hga, "1" = "K", "2" = "HSgrad", "3" = "attendCOL", "4" = "COLgrad", .default = "NA", .missing = "NA")) 

cpsasec_dfs$calculated_income <- as.factor(recode(cpsasec_dfs$htotval, "1" = "le15K", "2" = "15to25K", "3" = "25to35K", "4" = "35to50K", "5" = "50more", .default = "NA", missing = "NA"))

cpsasec_dfs$calculated_age <- as.factor(recode(cpsasec_dfs$a_age, "1" = "18to24", "2" = "25to34", "3" = "35to44", "4" = "45to54", "5" = "55to64", "6" = "65older", .default =  "NA", .missing = "NA"))

cpsasec_dfs$calculated_race <- as.factor(recode(cpsasec_dfs$prdtrace, "1" = "white", "2" = "black","3"="native","4"="asian","5"="pacific","6"="other" ,.default = "NA", .missing = "NA"))

cpsasec_dfs$iyear=as.factor(cpsasec_dfs$iyear)

    
#setting reference values
cpsasec_dfs$gender <- relevel(cpsasec_dfs$gender, ref = "female")
cpsasec_dfs$calculated_income <- relevel(cpsasec_dfs$calculated_income, ref = "50more")
cpsasec_dfs$calculated_education <- relevel(cpsasec_dfs$calculated_education, ref = "COLgrad")
cpsasec_dfs$calculated_age <- relevel(cpsasec_dfs$calculated_age, ref = "18to24")
cpsasec_dfs$calculated_race = relevel(cpsasec_dfs$calculated_race, ref = "white")

```

```{r, , include=FALSE, echo=FALSE}
table(cpsasec_dfs$ownstock)
cpsasec_dfs.bothsampling <- ovun.sample(ownstock ~  ., data = cpsasec_dfs, method = "under",  N=177296, seed = 1)$data
table(cpsasec_dfs.bothsampling$ownstock)

glm.ps.cps <- glm(ownstock ~ gender + calculated_education + calculated_income + calculated_age + calculated_race, family = binomial(link=logit),  data=cpsasec_dfs.bothsampling )
summary(glm.ps.cps)

#---Attach	the	predicted	propensity	score	to	the	datafile
cpsasec_dfs.bothsampling$psvalue.cps	<- predict(glm.ps.cps,	type	= "response")
hist(cpsasec_dfs.bothsampling$psvalue.cps,main="Histogram for Propensity of Stock Ownership", 
     xlab="Propensity Score", 
     border="black", 
     col="grey")
```


```{r, , include=FALSE, echo=FALSE}
#BRFSS
brfss_df13 <- 
    readRDS( file.path( path.expand( "~" ) , "BRFSS" , "2013 main.rds" ) )
brfss_df14 <- 
    readRDS( file.path( path.expand( "~" ) , "BRFSS" , "2014 main.rds" ) )
brfss_df15 <- 
    readRDS( file.path( path.expand( "~" ) , "BRFSS" , "2015 main.rds" ) )
brfss_df16 <- 
    readRDS( file.path( path.expand( "~" ) , "BRFSS" , "2016 main.rds" ) )
brfss_df17 <- 
    readRDS( file.path( path.expand( "~" ) , "BRFSS" , "2017 main.rds" ) )


brfss_df13$xpsu=as.integer(brfss_df13$xpsu)
brfss_df14$xpsu=as.integer(brfss_df14$xpsu)
brfss_df15$xpsu=as.integer(brfss_df15$xpsu)
brfss_df16$xpsu=as.integer(brfss_df16$xpsu)
brfss_df17$xpsu=as.integer(brfss_df17$xpsu)

### Create date
brfss_df13$iyear <- rep(2013,nrow(brfss_df13)) # make new column
brfss_df13$Date <- as.Date( paste(brfss_df13$iyear, brfss_df13$imonth , brfss_df13$iday , sep = "." )  , format = "%Y.%m.%d" )
brfss_df13=brfss_df13[order(as.Date(brfss_df13$Date, format="%d/%m/%Y")),]

brfss_df14$iyear <- rep(2014,nrow(brfss_df14)) # make new column
brfss_df14$Date <- as.Date( paste(brfss_df14$iyear, brfss_df14$imonth , brfss_df14$iday , sep = "." )  , format = "%Y.%m.%d" )
brfss_df14=brfss_df14[order(as.Date(brfss_df14$Date, format="%d/%m/%Y")),]

brfss_df15$iyear <- rep(2015,nrow(brfss_df15)) # make new column
brfss_df15$Date <- as.Date( paste(brfss_df15$iyear, brfss_df15$imonth , brfss_df15$iday , sep = "." )  , format = "%Y.%m.%d" )
brfss_df15=brfss_df15[order(as.Date(brfss_df15$Date, format="%d/%m/%Y")),]

brfss_df16$iyear <- rep(2016,nrow(brfss_df16)) # make new column
brfss_df16$Date <- as.Date( paste(brfss_df16$iyear, brfss_df16$imonth , brfss_df16$iday , sep = "." )  , format = "%Y.%m.%d" )
brfss_df16=brfss_df16[order(as.Date(brfss_df16$Date, format="%d/%m/%Y")),]

brfss_df17$iyear <- rep(2017,nrow(brfss_df17)) # make new column
brfss_df17$Date <- as.Date( paste(brfss_df17$iyear, brfss_df17$imonth , brfss_df17$iday , sep = "." )  , format = "%Y.%m.%d" )
brfss_df17=brfss_df17[order(as.Date(brfss_df17$Date, format="%d/%m/%Y")),]


brfss_df1314 <- intersect(colnames(brfss_df13), colnames(brfss_df14))

brfss_df1314 <- rbind(brfss_df13[ , brfss_df1314], brfss_df14[ , brfss_df1314])

brfss_df131415 <- intersect(colnames(brfss_df1314), colnames(brfss_df15))

brfss_df131415 <- rbind(brfss_df1314[ , brfss_df131415], brfss_df15[ , brfss_df131415])

brfss_df13141516 <- intersect(colnames(brfss_df131415), colnames(brfss_df16))

brfss_df13141516 <- rbind(brfss_df131415[ , brfss_df13141516], brfss_df16[ , brfss_df13141516])

brfss_df1314151617 <- intersect(colnames(brfss_df13141516), colnames(brfss_df17))

brfss_df1314151617 <- rbind(brfss_df13141516[ , brfss_df1314151617], brfss_df17[ , brfss_df1314151617])

brfss_df1314151617$state_name=factor(brfss_df1314151617$xstate , levels = 
                    c(1, 2, 4, 5, 6, 8, 9, 10, 11, 12, 13, 15, 16, 17, 18, 19, 20, 
                    21, 22, 23, 24, 25, 26, 27, 28, 29, 30, 31, 32, 33, 34, 35, 36, 
                    37, 38, 39, 40, 41, 42, 44, 45, 46, 47, 48, 49, 50, 51, 53, 54, 
                    55, 56) ,
                    
                labels = 
                    c("ALABAMA", "ALASKA", "ARIZONA", "ARKANSAS", "CALIFORNIA", 
                    "COLORADO", "CONNECTICUT", "DELAWARE", "DISTRICT OF COLUMBIA", 
                    "FLORIDA", "GEORGIA", "HAWAII", "IDAHO", "ILLINOIS", "INDIANA",
                    "IOWA", "KANSAS", "KENTUCKY", "LOUISIANA", "MAINE", "MARYLAND",
                    "MASSACHUSETTS", "MICHIGAN", "MINNESOTA", "MISSISSIPPI", 
                    "MISSOURI", "MONTANA", "NEBRASKA", "NEVADA", "NEW HAMPSHIRE",
                    "NEW JERSEY", "NEW MEXICO", "NEW YORK", "NORTH CAROLINA", 
                    "NORTH DAKOTA", "OHIO", "OKLAHOMA", "OREGON", "PENNSYLVANIA",
                    "RHODE ISLAND", "SOUTH CAROLINA", "SOUTH DAKOTA", "TENNESSEE",
                    "TEXAS", "UTAH", "VERMONT", "VIRGINIA", "WASHINGTON",
                    "WEST VIRGINIA", "WISCONSIN", "WYOMING"))

brfss_df= brfss_df1314151617 %>% select(state_name,	fmonth,	imonth,	iday,	iyear,	xpsu,	marital,	employ1,	sex,	xststr,	xstrwt,	xllcpwt,	xprace1,	xagexg,	xeducag,	xincomg,	Date, lsatisfy, drnkany5, cvdcrhd4,  xdrdxar1, diabete3, cvdstrk3, asthma3, chccopd1, addepev2,chcocncr, hlthpln1)

```

```{r, include=FALSE, echo=FALSE}
#DATA cleaning:

brfss_df$well_being = ifelse( brfss_df$lsatisfy %in% 1:4 , as.numeric( brfss_df$lsatisfy <= 2 ) , NA )

brfss_df$alcohol_consumption = ifelse(brfss_df$drnkany5==1,1,ifelse(brfss_df$drnkany5==2,0,NA))

brfss_df$cancer = ifelse( brfss_df$chcocncr ==1 , 1, ifelse(brfss_df$chcocncr==2, 0, NA))

brfss_df$health_insured = ifelse(brfss_df$hlthpln1 ==1 , 1, ifelse(brfss_df$hlthpln1==2, 0, NA))

brfss_df$heart_disease = ifelse( brfss_df$cvdcrhd4 ==1 , 1, ifelse(brfss_df$cvdcrhd4==2, 0, NA))

brfss_df$arthritis = ifelse( brfss_df$xdrdxar1 ==1 , 1, ifelse(brfss_df$xdrdxar1==2, 0, NA))

brfss_df$diabetes = ifelse( brfss_df$diabete3 ==1 , 1, ifelse(brfss_df$diabete3==2, 0, NA))

brfss_df$stroke = ifelse( brfss_df$cvdstrk3 ==1 , 1, ifelse(brfss_df$cvdstrk3==2, 0, NA))

brfss_df$asthma = ifelse( brfss_df$asthma3 ==1 , 1, ifelse(brfss_df$asthma3==2, 0, NA))

brfss_df$bronchitis = ifelse( brfss_df$chccopd1 ==1 , 1, ifelse(brfss_df$chccopd1==2, 0, NA))

brfss_df$depression = ifelse( brfss_df$addepev2 ==1 , 1, ifelse(brfss_df$addepev2==2, 0, NA))


#demographics
brfss_df$gender= ifelse( brfss_df$sex ==1 , 1, ifelse(brfss_df$sex==2, 0, NA))

brfss_df$marital_status <-  ifelse( brfss_df$marital ==1 , 1, ifelse(brfss_df$marital==2, 2,ifelse(brfss_df$marital==3,3,ifelse(brfss_df$marital==4,4,ifelse(brfss_df$marital==5,5,ifelse(brfss_df$marital==6,6, NA))))))

brfss_df$employment_status <- ifelse( brfss_df$employ1 ==1 , 1, ifelse(brfss_df$employ1==2, 2,ifelse(brfss_df$employ1==3,3,ifelse(brfss_df$employ1==4,4,ifelse(brfss_df$employ1==5,5,ifelse(brfss_df$employ1==6,6,ifelse(brfss_df$employ1==7,7,ifelse(brfss_df$employ1==8,8, NA))))))))

brfss_df$calculated_race <- ifelse( brfss_df$xprace1 ==1 , 1, ifelse(brfss_df$xprace1==2, 2,ifelse(brfss_df$xprace1==3,3,ifelse(brfss_df$xprace1==4,4,ifelse(brfss_df$xprace1==5,5,ifelse(brfss_df$xprace1==6,6,NA))))))

brfss_df$calculated_education <-ifelse( brfss_df$xeducag ==1 , 1, ifelse(brfss_df$xeducag==2, 2,ifelse(brfss_df$xeducag==3,3,ifelse(brfss_df$xeducag==4,4, NA))))

brfss_df$calculated_income <- ifelse( brfss_df$xincomg ==1 , 1, ifelse(brfss_df$xincomg==2, 2,ifelse(brfss_df$xincomg==3,3,ifelse(brfss_df$xincomg==4,4,ifelse(brfss_df$xincomg==5,5, NA)))))

brfss_df$calculated_age <- ifelse( brfss_df$xagexg ==1 , 1, ifelse(brfss_df$xagexg==2, 2,ifelse(brfss_df$xagexg==3,3,ifelse(brfss_df$xagexg==4,4,ifelse(brfss_df$xagexg==5,5,ifelse(brfss_df$xagexg==6,6,NA))))))

brfss_df=na.omit(brfss_df)
colSums(is.na(brfss_df))
```

```{r , include=FALSE, echo=FALSE}
getSymbols(c('^VIX', 'SPY','^DJI','^RUT'),from = '2013-01-01', to = '2017-12-30')

dat <- merge(VIX, SPY, all=TRUE)
dat$VIX.weeklyrets <- periodReturn(VIX,period='weekly',subset='2013-01-01::2017-12-30', type = 'log', leading = TRUE)
dat$VIX.monthlyrets <- periodReturn(VIX,period='monthly',subset='2013-01-01::2017-12-30', type = 'log', leading = TRUE)
dat$VIX.rets <- Delt(VIX[,6], type = 'log')
dat$SPY.rets <- Delt(SPY[,6], type = 'log')
dat$DJI.rets <- Delt(DJI[,6], type = 'log')
dat$RUT.rets <- Delt(RUT[,6], type = 'log')
# dat <- na.omit(dat)

dat_sentiment=data.frame(dat)
dat_sentiment$Date=rownames(dat_sentiment)
class(dat_sentiment$Date)

dat_sentiment=dat_sentiment[order(as.Date(dat_sentiment$Date, format="%Y.%m.%d")),]
dat_sentiment$date <- dat_sentiment$Date
dat_sentiment <- dat_sentiment %>% select(date, everything())
class(dat_sentiment$date)
dat_sentiment$date=as.Date(dat_sentiment$date, "%Y-%m-%d")

dat_sentiment$monthlag.VIX.Adjusted <- Lag(dat_sentiment$VIX.Adjusted, +20)


brfss_df$date=as.Date(brfss_df$Date, "%Y-%m-%d")


BRFSS_FINANCE_data= merge(dat_sentiment,brfss_df, by="date")

BRFSS_FINANCE_data= BRFSS_FINANCE_data %>% select(imonth,	iyear,	xpsu,	gender,	marital_status,	employment_status,	xststr,	xstrwt,	xllcpwt,	calculated_race,	calculated_education,	calculated_income,	calculated_age,	date, VIX.Adjusted, SPY.Adjusted, SPY.rets, VIX.rets, DJI.rets, RUT.rets, VIX.weeklyrets, VIX.monthlyrets,monthlag.VIX.Adjusted,  well_being, alcohol_consumption,cancer, health_insured, heart_disease, arthritis, diabetes, stroke, asthma, bronchitis, depression, state_name)

# X30returns,halfmonthreturns, X1weekreturns, VIX.rets, vix30returns, vixhalfmonthreturns, vix1weekreturns

#getting rid of duplicates
BRFSS_FINANCE_data=BRFSS_FINANCE_data[order(BRFSS_FINANCE_data$xpsu),]
BRFSS_FINANCE_data=BRFSS_FINANCE_data %>% distinct(xpsu, .keep_all = TRUE)

# colSums(is.na(BRFSS_FINANCE_data))
# 
# BRFSS_FINANCE_data=na.omit(BRFSS_FINANCE_data)

```


```{r , include=FALSE, echo=FALSE}
#DEMOGRAPHICS

BRFSS_FINANCE_data$gender=as.factor(recode(BRFSS_FINANCE_data$gender,'1'='male','0'='female',.missing='NA',.default = 'NA'))

BRFSS_FINANCE_data$marital_status <- as.factor(recode(BRFSS_FINANCE_data$marital_status, "1" = "married", "2" = "divorced","3"="widowed","4"="separated","5"="nevermarried","6"="membermarriedcoup",.default = "NA", .missing = "NA"))

BRFSS_FINANCE_data$employment_status <- as.factor(recode(BRFSS_FINANCE_data$employment_status, "1" = "wagesemployed", "2" = "selfemployed","3"="noworkmore1","4"="noworkless1","5"="homemaker","6"="student","7"="retired","8"="unable",.default = "NA", .missing = "NA"))

BRFSS_FINANCE_data$calculated_race <- as.factor(recode(BRFSS_FINANCE_data$calculated_race, "1" = "white", "2" = "black","3"="native","4"="asian","5"="pacific","6"="other" ,.default = "NA", .missing = "NA"))

BRFSS_FINANCE_data$calculated_education <- as.factor(recode(BRFSS_FINANCE_data$calculated_education, "1" = "K", "2" = "HSgrad", "3" = "attendCOL", "4" = "COLgrad", .default = "NA", .missing = "NA")) 

BRFSS_FINANCE_data$calculated_income <- as.factor(recode(BRFSS_FINANCE_data$calculated_income, "1" = "le15K", "2" = "15to25K", "3" = "25to35K", "4" = "35to50K", "5" = "50more", .default = "NA", missing = "NA"))

BRFSS_FINANCE_data$calculated_age <- as.factor(recode(BRFSS_FINANCE_data$calculated_age, "1" = "18to24", "2" = "25to34", "3" = "35to44", "4" = "45to54", "5" = "55to64", "6" = "65older", .default =  "NA", .missing = "NA"))

#setting reference values
BRFSS_FINANCE_data$gender <- relevel(BRFSS_FINANCE_data$gender, ref = "female")
BRFSS_FINANCE_data$calculated_income <- relevel(BRFSS_FINANCE_data$calculated_income, ref = "50more")
BRFSS_FINANCE_data$calculated_education <- relevel(BRFSS_FINANCE_data$calculated_education, ref = "COLgrad")
BRFSS_FINANCE_data$calculated_age <- relevel(BRFSS_FINANCE_data$calculated_age, ref = "18to24")
BRFSS_FINANCE_data$marital_status = relevel(BRFSS_FINANCE_data$marital_status, ref = "married")
BRFSS_FINANCE_data$employment_status = relevel(BRFSS_FINANCE_data$employment_status, ref = "wagesemployed")
BRFSS_FINANCE_data$calculated_race = relevel(BRFSS_FINANCE_data$calculated_race, ref = "white")

#year and state dummies
BRFSS_FINANCE_data$iyear = as.factor(BRFSS_FINANCE_data$iyear)
BRFSS_FINANCE_data$imonth = as.factor(BRFSS_FINANCE_data$imonth)
BRFSS_FINANCE_data$state_name=tolower(BRFSS_FINANCE_data$state_name)

#state-specific-time trends
BRFSS_FINANCE_data$yearmonth=as.yearmon( paste(BRFSS_FINANCE_data$iyear, BRFSS_FINANCE_data$imonth, sep = "." )  , format = "%Y.%m" )
BRFSS_FINANCE_data$yearmonth = as.factor(BRFSS_FINANCE_data$yearmonth)
BRFSS_FINANCE_data.drinking=BRFSS_FINANCE_data
```


```{r, , include=FALSE, echo=FALSE}
hist(cpsasec_dfs.bothsampling$psvalue.cps)

merge.data.brfss.cps= merge(BRFSS_FINANCE_data,cpsasec_dfs.bothsampling, by=c("gender" , "calculated_education" , "calculated_income" , "calculated_age" , "calculated_race"))
merge.data.brfss.cps=merge.data.brfss.cps[order(merge.data.brfss.cps$xpsu),]
merge.data.brfss.cps=merge.data.brfss.cps %>% distinct(xpsu, .keep_all = TRUE)

hist(merge.data.brfss.cps$psvalue.cps)
table(merge.data.brfss.cps$ownstock)

BRFSS_FINANCE_data <- merge.data.brfss.cps %>% select(-VIX.weeklyrets, - VIX.monthlyrets, - monthlag.VIX.Adjusted)
BRFSS_FINANCE_data.weekly <- merge.data.brfss.cps %>% select(- VIX.monthlyrets, - monthlag.VIX.Adjusted)
BRFSS_FINANCE_data.monthly <- merge.data.brfss.cps %>% select(- VIX.weeklyrets, - monthlag.VIX.Adjusted)

BRFSS_FINANCE_data<-na.omit(BRFSS_FINANCE_data)
BRFSS_FINANCE_data.weekly<-na.omit(BRFSS_FINANCE_data.weekly)
BRFSS_FINANCE_data.monthly<-na.omit(BRFSS_FINANCE_data.monthly)



brfss.drinking=merge.data.brfss.cps %>% select(-VIX.weeklyrets, - VIX.monthlyrets)
brfss.drinking <- na.omit(brfss.drinking)


mean(BRFSS_FINANCE_data$psvalue.cps)
median(BRFSS_FINANCE_data$psvalue.cps)
min(BRFSS_FINANCE_data$psvalue.cps)
max(BRFSS_FINANCE_data$psvalue.cps)

# BRFSS_FINANCE_data$l1.VIX.Adjusted <- Lag(BRFSS_FINANCE_data$VIX.Adjusted, +1)
# BRFSS_FINANCE_data$l2.VIX.Adjusted <- Lag(BRFSS_FINANCE_data$VIX.Adjusted, +2)
# BRFSS_FINANCE_data$l3.VIX.Adjusted <- Lag(BRFSS_FINANCE_data$VIX.Adjusted, +3)
# BRFSS_FINANCE_data$l4.VIX.Adjusted <- Lag(BRFSS_FINANCE_data$VIX.Adjusted, +4)
# BRFSS_FINANCE_data$l5.VIX.Adjusted <- Lag(BRFSS_FINANCE_data$VIX.Adjusted, +5)
# 
# BRFSS_FINANCE_data$l1.SPY.rets <- Lag(BRFSS_FINANCE_data$SPY.rets, +1)
# BRFSS_FINANCE_data$l2.SPY.rets <- Lag(BRFSS_FINANCE_data$SPY.rets, +2)
# BRFSS_FINANCE_data$l3.SPY.rets <- Lag(BRFSS_FINANCE_data$SPY.rets, +3)
# BRFSS_FINANCE_data$l4.SPY.rets <- Lag(BRFSS_FINANCE_data$SPY.rets, +4)
# BRFSS_FINANCE_data$l5.SPY.rets <- Lag(BRFSS_FINANCE_data$SPY.rets, +5)
# 
# BRFSS_FINANCE_data$l1.DJI.rets <- Lag(BRFSS_FINANCE_data$DJI.rets, +1)
# BRFSS_FINANCE_data$l2.DJI.rets <- Lag(BRFSS_FINANCE_data$DJI.rets, +2)
# BRFSS_FINANCE_data$l3.DJI.rets <- Lag(BRFSS_FINANCE_data$DJI.rets, +3)
# BRFSS_FINANCE_data$l4.DJI.rets <- Lag(BRFSS_FINANCE_data$DJI.rets, +4)
# BRFSS_FINANCE_data$l5.DJI.rets <- Lag(BRFSS_FINANCE_data$DJI.rets, +5)
# 
# BRFSS_FINANCE_data$l1.RUT.rets <- Lag(BRFSS_FINANCE_data$RUT.rets, +1)
# BRFSS_FINANCE_data$l2.RUT.rets <- Lag(BRFSS_FINANCE_data$RUT.rets, +2)
# BRFSS_FINANCE_data$l3.RUT.rets <- Lag(BRFSS_FINANCE_data$RUT.rets, +3)
# BRFSS_FINANCE_data$l4.RUT.rets <- Lag(BRFSS_FINANCE_data$RUT.rets, +4)
# BRFSS_FINANCE_data$l5.RUT.rets <- Lag(BRFSS_FINANCE_data$RUT.rets, +5)

#Random test
t.test(BRFSS_FINANCE_data$VIX.rets, BRFSS_FINANCE_data$SPY.rets)
t.test(BRFSS_FINANCE_data$VIX.rets, BRFSS_FINANCE_data$RUT.rets)
t.test(BRFSS_FINANCE_data$VIX.rets, BRFSS_FINANCE_data$DJI.rets)

cor(BRFSS_FINANCE_data$VIX.rets, BRFSS_FINANCE_data$SPY.rets)
cor(BRFSS_FINANCE_data$VIX.rets, BRFSS_FINANCE_data$RUT.rets)
cor(BRFSS_FINANCE_data$VIX.rets, BRFSS_FINANCE_data$DJI.rets)

```


```{r, include=FALSE, echo=FALSE}
#Starting Analysis
#All Population
table(BRFSS_FINANCE_data$well_being)
data_balanced_both <- ovun.sample(well_being ~  ., data = BRFSS_FINANCE_data, method = "both", p=0.5,  N=7516, seed = 265)$data

table(BRFSS_FINANCE_data.weekly$well_being)
data_balanced_over.weekly <- ovun.sample(well_being ~ ., data = BRFSS_FINANCE_data.weekly, method = "over", N=2574, seed = 1)$data 
data_balanced_both.weekly <- ovun.sample(well_being ~  ., data = BRFSS_FINANCE_data.weekly, method = "both", p=0.5,  N=1397, seed = 44)$data #13, 19, 42 (GOOD), 44 (BESTest), 53, 56, 58 (GOOD), 64 (OK), 69 (GOOD), 74 (GOOD), 92 (GOOD), 

table(BRFSS_FINANCE_data.monthly$well_being)
data_balanced_over.monthly <- ovun.sample(well_being ~  ., data = BRFSS_FINANCE_data.monthly, method = "over", N=346, seed = 1)$data 
data_balanced_both.monthly <- ovun.sample(well_being ~  ., data = BRFSS_FINANCE_data.monthly, method = "both", p=0.5,  N=181, seed = 1)$data 
```

```{r, include=FALSE, echo=FALSE}
#GOF 
#Overall Population - both sampling
#VIX.rets
glm.vixrets.both.nochronic = glm(well_being ~ VIX.rets + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x, family=binomial(link=logit),  weights=xstrwt, data = data_balanced_both)
sum.glm.vixrets.both.nochronic=summary(glm.vixrets.both.nochronic)
sum.glm.vixrets.both.nochronic

glm.vixrets.both.nointeraction = glm(well_being ~ VIX.rets + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x + health_insured + heart_disease + arthritis + stroke + asthma + bronchitis + depression + cancer + diabetes, family=binomial(link=logit),  weights=xstrwt, data = data_balanced_both)
sum.glm.vixrets.both.nointeraction=summary(glm.vixrets.both.nointeraction)
sum.glm.vixrets.both.nointeraction

glm.vixrets.both = glm(well_being ~ VIX.rets*psvalue.cps + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x + health_insured + heart_disease + arthritis + stroke + asthma + bronchitis + depression + cancer + diabetes, family=binomial(link=logit),  weights=xstrwt, data = data_balanced_both)
sum.glm.vixrets.both=summary(glm.vixrets.both)
sum.glm.vixrets.both

partial.effect.both.mean <- coef(glm.vixrets.both)[2] + coef(glm.vixrets.both)[63] * mean(data_balanced_both$psvalue.cps)
partial.effect.both.med <- coef(glm.vixrets.both)[2] + coef(glm.vixrets.both)[63] * median(data_balanced_both$psvalue.cps)
partial.effect.both.min <- coef(glm.vixrets.both)[2] + coef(glm.vixrets.both)[63] * min(data_balanced_both$psvalue.cps)
partial.effect.both.max <- coef(glm.vixrets.both)[2] + coef(glm.vixrets.both)[63] * max(data_balanced_both$psvalue.cps)

#VIX.weeklyrets
glm.vixrets.both.nochronic.weekly = glm(well_being ~ VIX.weeklyrets + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x, family=binomial(link=logit),  weights=xstrwt, data = data_balanced_both.weekly)
sum.glm.vixrets.both.nochronic.weekly=summary(glm.vixrets.both.nochronic.weekly)
sum.glm.vixrets.both.nochronic.weekly

glm.vixrets.both.nointeraction.weekly = glm(well_being ~ VIX.weeklyrets + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x + health_insured + heart_disease + arthritis + stroke + asthma + bronchitis + depression + cancer + diabetes, family=binomial(link=logit),  weights=xstrwt, data = data_balanced_both.weekly)
sum.glm.vixrets.both.nointeraction.weekly=summary(glm.vixrets.both.nointeraction.weekly)
sum.glm.vixrets.both.nointeraction.weekly

glm.vixrets.both.weekly = glm(well_being ~ VIX.weeklyrets*psvalue.cps + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x + health_insured + heart_disease + arthritis + stroke + asthma + bronchitis + depression + cancer + diabetes, family=binomial(link=logit),  weights=xstrwt, data = data_balanced_both.weekly)
sum.glm.vixrets.both.weekly=summary(glm.vixrets.both.weekly)
sum.glm.vixrets.both.weekly

partial.effect.both.mean.weekly <- coef(glm.vixrets.both.weekly)[2] + coef(glm.vixrets.both.weekly)[61] * mean(data_balanced_both$psvalue.cps)
partial.effect.both.med.weekly <- coef(glm.vixrets.both.weekly)[2] + coef(glm.vixrets.both.weekly)[61] * median(data_balanced_both$psvalue.cps)
partial.effect.both.min.weekly <- coef(glm.vixrets.both.weekly)[2] + coef(glm.vixrets.both.weekly)[61] * min(data_balanced_both$psvalue.cps)
partial.effect.both.max.weekly <- coef(glm.vixrets.both.weekly)[2] + coef(glm.vixrets.both.weekly)[61] * max(data_balanced_both$psvalue.cps)

latex_general_vixrets.both=stargazer(glm.vixrets.both.nochronic,glm.vixrets.both.nointeraction,glm.vixrets.both, type="latex",omit = c("imonth","state_name.x","iyear.x"),omit.labels =c("Month FE","State FE", "Year FE"), title="Regression Results for Well-Being and VIX - Under and Over Sampling Combination")

latex_general_vixrets.both.weekly=stargazer(glm.vixrets.both.nochronic.weekly,glm.vixrets.both.nointeraction.weekly,glm.vixrets.both.weekly, type="latex",omit = c("imonth","state_name.x","iyear.x"),omit.labels =c("Month FE","State FE", "Year FE"), omit.stat = c("rsq", "f"), title="Regression Results for Well-Being and VIX - Under and Over Sampling Combination")

```


```{r, include=FALSE, echo=FALSE}
#alcohol_consumption model
brfss.drinking <- na.omit(brfss.drinking)
table(brfss.drinking$alcohol_consumption)

data_balanced_both.drinking <- ovun.sample(alcohol_consumption ~  ., data = brfss.drinking, method = "both", p=0.5,  N=7376, seed = 2)$data #seed is 29

#based on previous literature, use log(VIX.Adjusted)
model.glm.drinking.nochronic= glm(alcohol_consumption ~  log(monthlag.VIX.Adjusted) + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x , family=binomial(link=logit), weights=xstrwt, data  = data_balanced_both.drinking)
sum.model.glm.drinking.nochronic=summary(model.glm.drinking.nochronic)
sum.model.glm.drinking.nochronic

model.glm.drinking.nointeraction= glm(alcohol_consumption ~  log(monthlag.VIX.Adjusted) + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x + health_insured + heart_disease + arthritis + stroke + asthma + bronchitis + depression + cancer + diabetes, family=binomial(link=logit), weights=xstrwt, data  = data_balanced_both.drinking)
sum.model.glm.drinking.nointeraction=summary(model.glm.drinking.nointeraction)
sum.model.glm.drinking.nointeraction

model.glm.drinking= glm(alcohol_consumption ~  log(monthlag.VIX.Adjusted)*psvalue.cps + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x + health_insured + heart_disease + arthritis + stroke + asthma + bronchitis + depression + cancer + diabetes, family=binomial(link=logit), weights=xstrwt, data  = data_balanced_both.drinking)
sum.model.glm.drinking=summary(model.glm.drinking)
sum.model.glm.drinking

partial.effect.drinking.mean <- coef(model.glm.drinking)[2] + coef(model.glm.drinking)[63] * mean(data_balanced_both.drinking$psvalue.cps)
partial.effect.drinking.med <- coef(model.glm.drinking)[2] + coef(model.glm.drinking)[63] * median(data_balanced_both.drinking$psvalue.cps)
partial.effect.drinking.min <- coef(model.glm.drinking)[2] + coef(model.glm.drinking)[63] * min(data_balanced_both.drinking$psvalue.cps)
partial.effect.drinking.max <- coef(model.glm.drinking)[2] + coef(model.glm.drinking)[63] * max(data_balanced_both.drinking$psvalue.cps)

latex_general_drinking.full=stargazer(model.glm.drinking.nochronic, model.glm.drinking.nointeraction, model.glm.drinking, type="latex",omit = c("imonth","state_name.x","iyear.x"),omit.labels =c("Month FE","State FE", "Year FE"), omit.stat = c("rsq", "f"), title="Regression Results for Alcohol Consumption and Monthly VIX - Under and Over Sampling Combination")

```

```{r, include=FALSE, echo=FALSE}
#robustness check
#logit
table(data_balanced_both$well_being)
# Create Training Data
input_ones <- data_balanced_both[which(data_balanced_both$well_being == 1), ]  # all 1's
input_zeros <- data_balanced_both[which(data_balanced_both$well_being == 0), ]  # all 0's
set.seed(150)  # for repeatability of samples
input_zeros_training_rows <- sample(1:nrow(input_zeros), 0.7*nrow(input_zeros)) 
input_ones_training_rows <- sample(1:nrow(input_ones), 0.7*nrow(input_zeros))  
 
training_ones <- input_ones[input_ones_training_rows, ]  
training_zeros <- input_zeros[input_zeros_training_rows, ]
trainingData <- rbind(training_ones, training_zeros)  # row bind the 1's and 0's 
table(trainingData$well_being)

# Create Test Data
test_ones <- input_ones[-input_ones_training_rows, ]
test_zeros <- input_zeros[-input_zeros_training_rows, ]
testData <- rbind(test_ones, test_zeros)  # row bind the 1's and 0's 
table(testData$well_being)
```

```{r , include=FALSE, echo=FALSE, eval=FALSE}
descrips_financial_brfss_lag =
  BRFSS_FINANCE_data %>%
  dplyr::select(well_being, calculated_age ,gender,calculated_income,calculated_education,marital_status,employment_status,calculated_race,state_name.x)

descriptive_statisics <- descrips_financial_brfss_lag %>%
  # build base summary table
  tbl_summary(
    by = well_being,
    # change variable labels
    label = list(well_being ~ "Well-Being",
                 calculated_age ~ "Age",
                 gender ~ "Gender",
                 calculated_income ~ "Income",
                 calculated_education ~ "Education",
                 marital_status ~ "Marital Status",
                 calculated_race ~ "Race",
                 employment_status ~ "Employment Status",
                 state_name.x ~ "State Name"),
    # change statistics printed in table
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} / {N} ({p}%)"),
  ) %>%
  # add statistic labels
  add_stat_label() %>%
  # bold variable labels, italicize levels
  bold_labels() %>%
  italicize_levels() %>%
   gt() %>%
  tab_header(
    title = md("Descriptive Statistics"),
    subtitle = md("By Well-Being")
  ) %>%
  as_latex()

descriptive_statisics %>%
  as.character() %>%
  cat()


```

```{r , include=FALSE, echo=FALSE, , eval=FALSE}
descriptive_statistics_binary =  BRFSS_FINANCE_data %>%
  dplyr::select(health_insured, heart_disease, arthritis, stroke,asthma, bronchitis,depression, cancer, diabetes  )

stargazer(descriptive_statistics_binary,  title="Descriptive statistics for Binary Variables", covariate.labels=c("Health Insured", "Heart Disease", "Arthritis","Stroke","Asthma", "Bronchitis", "Depression", "Cancer", "Diabetes"))

```

\abstract{We investigate how self-reported well-being varies with the volatility (VIX) from stock volatility of 500 large companies listed on stock exchanges in the United States (S\&P 500). Because stock prices represent the discounted value of the company’s future cash-flows, aggregated market indices, such as the S\&P 500, are forward looking in nature and thus provide leading economic indicators. By studying the relationship between individual well-being levels and anticipated market volatility, this research contributes to our understanding of the psychological impact of future economic uncertainty. Prior literature examining this topic typically relies on board economic indicators, such as the unemployment rate (Böckerman, Petri, and Ilmakunnas, 2009 and Brenner, M. Harvey, and Mooney, 1983), or directional changes in market indices, such as the Dow Jones Industrial Average (Cotti, Dunn, and Tefft, 2013). We propose that a volatility index, namely the VIX, offers a more robust empirical assessment of how financial market conditions affect mental and physical states of being and postulate that volatility clusters provide a better indication of market turmoil than simply looking at negative returns. After accounting for associations with the unemployment rate and stock ownership, we find positive relationship between the VIX and poor self-reported well-being. These results are consistent with additional models for robustness check in this analysis.}




# 1. Introduction

Modern finance theory postulates that asset prices in a competitive market reflect all available public information, so that in the limit, prices only move in response to news. Thus, a stock’s price is an average of the market’s sentiment regarding the future performance of the underlying company, and movements from this equilibrium price reflects the interpretation of new and exogenous information regarding the expected cash flows of the firm. At the index level, price changes reflect updates to the forecasted free cash-flows of a specific sector or the market as a whole and provides a forward-looking indicator of macroeconomic performance. For this study, which assess the impact of market conditions on well-being, we are concerned the aggregate level of uncertainty surrounding future price movements – the tightening and widening of confidence intervals around price forecasts, which is reflected through changes in the price of options contracts. As uncertainty increases, out-of-the-money call and put options become more expensive, i.e the implied likelihood of out-of-the-money options increases. Thus, aggregating these prices across the broader market provides a proxy for the general level of economic doubt, precariousness, and unease.

In order to capture this effect, similar papers have used this dataset to study the relationship between self-reported well-being variables and market indicators, such as the Dow Jones Industrial Average (DJIA) (Cotti, Dunn, and Tefft, 2013). We approach this question from a slightly different angle; rather that assessing the correlation between well-being and directional changes in market indices, we measure the relationship between well-being and the magnitude of \emph{anticipated} changes in market indices – market \emph{implied} volatility. We believe this is an improvement for several reasons. First, periods of market turmoil are characteristically marked by large price movements in both directions – volatility clusters. That is, large market declines may be followed by a large transitory rebound, which is then followed by another large decline. We propose that these transitory rebounds amidst a broader crisis do not provide psychological relief in equal proportion to the distress caused by a market fall of equal magnitude. Thus, our empirical model should capture the market’s level of uncertainty (implied volatility) rather than noisy directional price changes, to which a linear estimation model would be sensitive. Second, the perception of well-being, and the associated psychological response, is non-linear between the domains of gains and losses, as suggested by Prospect Theory (Kahneman and Tversky, 1979). To capture these non-linear effects, our model employs the widely cited volatility index, VIX, as the independent variable of interest, which aggregates executed near and next term out-of-the-money calls and puts across all S&P 500 options contracts. It is well documented that the VIX is asymmetrical in its response to changes in the underlying S\&P 500 index, rising more following downside volatility (Low, 2004). This asymmetrical response between the domains of gains and losses is similar to the emotional response predicted by Prospect Theory. 

In addition to providing an aggregate view of the market’s current short-term volatility expectations, the VIX has additional properties that are of interest for this study, namely its presence in the news media and widespread recognition as the market’s fear gauge. Thus, from a psychological perspective, the VIX carries a pre-established negative connotation and is often cited alongside the major market indices. With the VIX, we capture not only market risk and turmoil, but also a general sense of uncertainty and fear about future macroeconomic conditions, which, we propose, drives psychological and physical malaise.

Research demonstrates a significant, yet complicated, role for the news media in shaping economic perceptions. Now more than ever, we are confronted with 24-hour market coverage, push notifications on cell phones, and instant access to stock portfolios via mobile applications. Through this increasingly accessible and rapid media coverage, market signals reach a significant percentage of the general population and help shape sentiment regarding the economic outlook and confidence about one’s current and future socio-economic well-being (Procopio, Terrell, & Wu, 2010). Within this context, signals of increased uncertainty have a diminishing effect on one’s well-being, both economically and physically.

Because most Americans do not actively trade stocks or derivatives, market signals do not typically represent the realization of gains and losses to capital income. Instead, market changes represent changes in wealth since capital gains and losses are only realized upon the sale of an asset. Following prior research on this topic, we limit the income effect that would result from a change in macroeconomic conditions by controlling for unemployment and per capita personal income. Doing so allows us to more effectively capture the effects of market stress and uncertainty.

During the recent stock market crashes, Americans reported large declines in self-reported well-being (Deaton, 2011), exhibited increased symptoms of depression and poor mental well-being (McInerney, Mellor, & Nicholas, 2012), and experienced a spike in hospitalizations for psychological disorders (Engelberg & Parsons, 2013). Similarly, we anticipate significant and important changes to self-reported general well-being during periods of market turmoil and uncertainty. By using survey data from the Behavioral Risk Factor Surveillance System (BRFSS) between 2013 and 2017, we estimate the relationship between expected S&P 500 volatility and overall well-being.

From a sensory perspective, uncertainty plays a major role in how the nervous system interprets and perceives the physical world, and how external signals affect to the realm of the mind where we interpret information. Because we are only ever presented with an incomplete potion of the world at a given time, our sensory nervous system must make sense of noisy signals during the best of times. Therefore, it may be argued that a central purpose of the nervous system is to estimate the state of the world from noisy and incomplete data (Vilares and Kording, 2012). As signal noise increases during uncertain economic and financial periods, our physical and mental systems are abnormally stressed, causing both physical and mental harm.

To demonstrate this effect, key models of anxiety-related psychopathology describe anxiety and fear as having three key components: physiological (i.e., autonomic nervous system activation), cognitive (i.e., interpretation of the environmental and physiological stimuli) and behavioral (i.e., the selected response to the stimuli). Research in this field suggests that with respect to the experience of anxiety, the most threatening stimuli are consequentially threatening, in that threatening stimuli (e.g., losing money) are threatening as a function of anticipated consequences (e.g., decreased welfare and utility) and patterns of reinforcement, rather than being inherently threatening (Carleton, 2014). To frame this concept in an economic sense, increases in market uncertainty are threatening as a function of the anticipated consequences of a poor economy. It is this relationship, the correlation between anticipated economic conditions and contemporaneous well-being indicators, that we assess in this paper. From a methodological standpoint, we follow prior literature regarding additional control variables and find that our results demonstrate a robustness to demographic characteristics (e.g. gender, race), income and employment status, area and time fixed effects, and several additional factors that may influence survey responses. As far as we are aware, this is the first study to measure the relationship between individuals’ self-report well-being and the market’s fear gauge, VIX.

Using observational survey data from Behavioral Risk Factor Surveillance System (BRFSS), Current Population Survey (CPS) and Chicago Board Options Exchange (CBOE) Volatility Index (VIX) as the measure of expected stock market uncertainty from 2013 to 2017, this paper finds strong support for our hypothesis. Adjusting for demographics, health conditions, and different fixed effects for time and states, this study reveals that the VIX has a negative influence on well-being. More significantly, whether the respondent owns the stock or not, this impact is permanent. This effect is greater as the stock holding propensity is higher and lower as the stock holding propensity is smaller. In specific, additional increase in market volatility return decreases well-being by an expected log count of 0.457 for respondents who are least likely to own stock, 0.794 for respondents who have a fifty-fifty chance of owning stocks, and 1.21 for respondents who are most likely to own stock. This clearly shows that income effect is present in our model, where the effect of stock market is more prevalent to people who have steaks in the stock market. In addition, we also proposed a second model in this paper to support our main result. As one of our robustness checks, inspired by Sias 2017, we figure the relationship of market volatility on life satisfaction and well-being can be translated into many channels. If market volatility has an effect on individual well-beings, then it also should have an effect on behaviors, whether it is direct or indirect impact. Using the same dataset, this study finds that VIX has positive influence on impulse control, particularly on alcohol consumption. Specifically, additional increase in the natural log of market volatility increases the likelihood of consume alcohol by an expected log count of 0.487 for respondents who are least likely to own stock, 1.053 for respondents who have a fifty-fifty chance of owning stocks, and 1.556 for respondents who are most likely to own stock.

Our results have a range of major implications. First, the relationship between market volatility and control of impulses can help to explain the financial decisions of individuals. For example, a decrease in impulse control can help explain why when expected market volatility is strong, behavioral biases are more extreme (Kumar, 2009). In other words, understanding what drives investor decisions is essential to understanding behavioral finance. Our findings indicate that one of the channels connecting economic conditions to the actions of investors could be the relationship between anticipated market volatility and impulse control. Second, our findings shed lights into papers indicating the effect of anticipatory feelings (e.g., Lowenstein, 1987) on individual desires (and therefore behaviors). Caplin and Leahy (2001) demonstrate, for example, that adding sentiments to the utility function can help explain time inconsistency in preferences. We show that the effect of forward-looking volatility fits into the concept of anticipatory theory framework, in which the VIX acts as the anticipatory index and well-being is measure of utility. Third, our findings add on to the literature in regard to feedback models (e.g. Shiller, 2002). Specifically, as Engelberg and Parsons (2016) have pointed out, most behavioral finance work concentrates on how investor behavior affects markets and avoids opposite effect. As a result, our finding introduces a new connection of how markets influence investor behavior.

This study is not the first to investigate the relationship between market uncertainty and health indicators. In fact, our study is motivated by recent behavioral finance papers (Engelberg and Parsons 2016, Sias 2017, etc.). However, this study differs from previous studies in two major ways. First, this study is the first to use market volatility return as the main independent variable of interest. Previous studies used the natural log of market volatility, however, we finds that the market volatility returns are much more stationary and thus provide better and more robust results. Second, we link the effect of market volatility on well-being through the anticipatory theoretical framework and shows that our model clearly deviates from the traditional neo-classical model. In specific, we show that the VIX can be a natural anticipatory index, which can be built under the well-being utility function (Stevenson, Wolfers 2008). 

This paper proceeds as follow. Section 2 outlines the data and theoretical mechanisms of market uncertainty and individuals’ well-beings. Section 3 presents the empirical approach and descriptive statistics. Section 4 illustrates the methodologies used on the dataset and presents the results. This section also discusses goodness of fits of the estimates produced in this paper. Section 6 concludes the paper.


# 2. Data and Theoretical Framework

### 2.1 Data

Our measure of well-being outcome comes from the Behavioral Risk Factor Surveillance System (BRFSS), maintained by the Center for Disease Control and Prevention (CDC) to monitor health and behavioral risk in the U.S. Using the BRFSS data from 2013-2017, we construct the measure of each individual’s well-being using an indicator for general life-satisfaction in the BRFSS survey. This variable is recorded from the very first general health question under Health Status section, which asks, "In general, how satisfied are you with your life?" with the answers ranging from "Very satisfied"," Satisfied", "Dissatisfied" to "Very dissatisfied". We recode this variable to a binary outcome, in which "Very satisfied" and "Satisfied" are grouped to 1, and "Dissatisfied" and "Very dissatisfied" are grouped to 0. We get rid of respondents who answer, "Don’t know/Not Sure", "Refused", and missing data. Given that the BRFSS survey is not consistent with the questions and the variable codes, especially before 2013, it is very difficult to match all the same variables historically. In addition, each year's dataset asks different questions, which populate different variables. Depending on the year, the number of variables can range from 180 to 300. Therefore, we only match variables that are consistent over the entire sample periods. To further prepare the data, we only keep variables that are being analyzed in this study and eliminate all NAs from this subset. Moreover, we also get rid of observations from respondents that did not complete the survey. For demographic variables, given the nature of the BRFSS survey, all of them are categorical variables. Recoding for all variables including $age$, $gender$, $income$, $education-status$, $marital-status$, $employment-status$, $race$ are performed consistently, where missing data, "don't know," and "refused" are eliminated. For detail descriptions and recoding process, see Appendix. 

As discussed, our main variable of interests is the Chicago Board Options Exchange (CBOE) Volatility Index, or VIX. VIX has been referred to as the ‘investor fear gauge’ (Whaley (2000)), since high levels of VIX coincided with high degrees of market turmoil. The VIX index was originally designed to measure the market’s expectation of 30-day volatility implied by at-the-money S&P 100 Index (OEX Index) option prices. The popularity of the VIX Index made it soon become the premier benchmark for U.S. stock market volatility. The new VIX Index is taken from the S&P 500 Index (SPXSM). The idea was to base on the core index for U.S. equities, and estimates expected volatility by aggregating the weighted prices of SPX puts and calls over a wide range of strike prices. Daily adjusted returns for the VIX will used as the main variable of interest in this analysis. The VIX model for market volatility is presented in the main Panel.

Financial data is extracted from Yahoo Finance! Since stock market only has observations 5 out of 7 days of the week (except some holidays), matching the daily returns of the VIX to BRFSS historical survey data (which covers almost every day of the year including the weekends) resulted in the loss of a lot of observations. The return series for the VIX follows the return formula of $\frac{VIX_t-VIX_{t-1}}{VIX_{t-1}}$, where the VIX is the adjusted price for the implied volatility index, the VIX. We refer to this return series as the VIX in this paper. Performing the regression using the VIX return series instead of the adjusted price not only separates this paper to relevant papers but also gives a more robust and more stationary result. 

To better assess the effect of market volatility on reported well-being, we attempt to control for stock ownership in our analysis. The BRFSS survey unfortunately does not provide information on whether the respondents hold stocks, however, we were able to achieve this information from Current Population Survey (CPS). The CPS surveys, conducted monthly by the U.S. Bureau of the Census and the U.S. Bureau of Labor Statistics, are representative of the entire U.S. population. Although the CPS does not have daily data like the BRFSS, which makes it impossible to merge with the Financial data, it provides some similar characteristics in term of demographics to the BRFSS, and most importantly it has stockholding information, which it is incredibly useful for our analysis. CPS data is taken from 2013-2017 with the total observations of 704,345 observations. Suggested by previous literature (Kreinin et al 1959), we use some of the most significant demographic predictors of stock ownership, including age, gender, income, education-status, and race, to merge stockholding data from the CPS onto the BRFSS and the VIX. All of these variables from the CPS are recoded from continuous to categorical data to match with the BRFSS demographic variables. ^[See Section 4 for methodology.]


### 2.2 Theoretical Point of Views

#### Can measures of well-being be interpreted as utility?

In response to Stevenson, Wolfers (2008), "Economic growth and subjective Well-being: reassessing the Easterlin Paradox", Becker and Rayo assert that this paper not only provides convincing evidence that self-reported happiness and measures of life satisfaction are positively correlated with income, both in rich and poor countries, but also provides an interesting take on the relationship between utility and reported happiness or life satisfaction. They also conclude that, although there are grey areas in connecting self-reported life satisfaction to utility, it is quite acceptable to view happiness and life satisfaction as “noisy measures of utility.” However, while they agree that there are connections between the two dimensions, "reported happiness and life satisfaction are no more measures of utility than are other dimensions of well-being, such as health or consumption of material goods." Prior literature has shown that happiness levels are very consistent with reported happiness, whether it is self-reported or reported by a trusted third party. Studies have also shown that we, as human beings, behave consistently with survey reports as well. For instance, we try to avoid bad situations that may result in a reduction in self-reported happiness.  
 
At the same time, Becker and Rayo suggest an alternative interpretation of life-satisfaction and happiness data, in which these measures can be interpreted as a commodity in the utility function. In many perspectives, Becker and Rayo implicitly assert that just like owning a house or being healthy, happiness and life-satisfaction indeed describe the same commodity in the utility function. In addition, they take an extra step to show that there are ways to use the consumer utility maximization theory to test whether well-being can be used as utility. In other words, it is possible to test whether happiness data is in fact a commodity in the utility function. While discussion on whether happiness is identical with utility still needs more research, it is important to accept that life-satisfaction and happiness constitute useful measures in consumer utility maximization theory.  
 
Given these reasons, while we do not necessarily take life-satisfaction or happiness as our measures of utility, we are convinced that using self-reported health could be a reasonable approach to measure utility, a commodity within a utility function, or at least a "noisy" measure of utility.  

#### Can the VIX be interpreted as anticipatory index?

In the neo classical model, people derive their utility from the consumption of physical goods, such as houses, food, health, leisure, etc. People also derive their utility from their beliefs. For instance, being excited for a trip or worrying about health are anticipatory emotions that are derived from beliefs; believing food taste better if it was made from quality ingredients is belief about consumables; avoiding negative information about oneself is belief about self, etc. There are some examples of belief-based preferences. Regarding consumption, people also have beliefs about their future consumption, which, importantly, can cause experienced utility in the present. This is called anticipatory utility. The belief of diminished future wealth or a more pessimistic future can cause pain now. Similarly, belief of a more optimistic future regarding employment status or wealth can bring contemporaneous enjoyment.  
 
VIX, or the Volatility Index was originally designed to measure the market’s expectation of 30-day volatility implied by at-the-money S&P 100 Index (OEX Index) option prices. Thus, by definition, it is a forward-looking index, and arguably one of the best measures to make prediction about the future economy. It is also a measure of uncertainty about the future, since large changes in asset prices generally result from contemporaneous uncertainty about the future health of the economy. Regarding the anticipatory model, considering the VIX as a future belief index is reasonable, and thus we hypothesize that changes in this index can influence present enjoyment or well-being. 
 
This concept was proven in 1987 when Loewenstein first started to look at time discounting, which is the first implication of the anticipatory model. This paper examines a different implication of the model, namely preferences for information. Anticipatory utility is built from beliefs about the future. This information can be expressed in the VIX, where it could provide necessary information about whether the economy is expected to be better or worse in the near future. In other words, just like anticipation from the kiss or from having to clean up the hamster cages in Loewenstein model, anticipation about the future economy should have similar mechanism. For instance, similar to anticipation from the kiss, anticipation from a bright future economy should have both consumption and anticipation benefits. Here, these benefits translate to a better total instantaneous well-being, where self-reported well-being at this current moment is the consumption utility. 

#### Neo-classical vs Anticipatory Theory?

Consider a $T-period$ model, where the agent decides in period $0$ when to consume. Let $x_t$ be consumption at time t. $0 \leq \delta \leq 1$ is the discount factor. Utility from consumption in period $t$ is $u(x_t)$, the total discount utility in neo-classical model of preferences overtime is presented by: $U_0(x)= \sum_{t=0}^T \delta^t u(x_t)$. In anticipatory model, utility from anticipation in period $t$ is $\alpha \sum_{i=t+1}^T \delta^{i-t}u(x_i)$, where $\alpha > 0$. The total instantaneous utility at $t$ is presented by: $u(x_t) + \alpha \sum_{i=t+1}^T \delta^{i-t}u(x_i)$. And the total anticipatory utility over consumption is: $U_t=\sum_{j=t}^T \delta^{j-t}[u(x_j)+\alpha \sum_{i=j+1}^T \delta^{i-j}u(x_i)]$. 
  
The Neo-classical model only gives credit to physical goods, from which one’s utility can derive. We clearly shy away from the neo-classical model, and by adopting the anticipatory model we assume the VIX to be the consumption for beliefs. Therefore, the effect of beliefs on instantaneous utility is presented by $u(x_t)$. In addition, to better present our hypothesis, we adapt an extended model by Caplin and Leahy (2001). This model is an add-on to the previous model regarding the role for information since it is difficult to talk about a future that is uncertain.  
 
There is uncertainty over the outcome of future events, and thus agents may not want to know information since it could alter the beliefs, which eventually results in different decision-makings. A simplified Caplin and Leahy model assumes two periods. The decision maker may receive information in period 1 and the outcome is realized in period 2. Let $x$ be random variable over support S and with distribution $f(x)$, the consumption utility over $x$ by decision maker is presented by $u(x)$. The decision maker also has anticipatory utility in period 1, $b(x)$, over beliefs of $x$, and $a(b(x))$ is anticipatory utility. If there is no time discounting, the total expected utility in period 1 is: $U_1=a(b(x)) +\sum_{x \in S} b(x)u(x)$. Without information, the decision maker has prior belief $f(x)$ and with information, the decision maker has posterior belief $g(x)$. The decision maker only wants information if: $E_{f_x}[a(g(x)) + \sum_{x \in S} g(x)u(x)]>a(f(x)) + \sum_{x \in S} f(x)u(x)$. Given the equation, the decision maker only wants information if and only if: $E_{f_x}[a(g(x))] > a(f(x))$. Under classical model, $a(b(x))$ does not exist, and thus with or without beliefs, the classical decision maker should not be affected. 
  
To summarize the model, under a simplified two-period setup, a classical agent is completely indifferent to receiving future information, whether the information is negative or positive. This implies that information on the VIX which is the forward-looking information should not influence the agent’s utility, or agent’s reported well-being under the classical model. In other words, even in our data where it is applicable for both stockholders and non-stockholders, the classical model would predict that that even stockholders who frequently receive market news should not be affected by this information. Our paper proves otherwise. 

#### Mechanism of Market Uncertainty and Individuals' Well-beings

We propose three different explanations for the impact of market uncertainty, recorded using the VIX, on individuals' well-beings. 

1. The first explanation is straight forward, with the effects of uncertainty explained through the mechanisms of the anticipatory theoretical framework. Market uncertainty imposes stress on life-satisfaction and behaviors, thus decreasing reported life-satisfaction. The effect of market volatility on different behavioral measures has been studied in many previous research efforts. For instance, Kalcheva, McLemore, Sias 2017 showed that the VIX has an prominent impact on different impulsive behaviors such as drinking and smoking; Schwandt 2014 shows evidence of market volatility under the S&P500 on physical health, mental health and survival rates; Schwandt, Hannes 2018 updated on Schwandt 2014 to estimate the same effect during boom and bust markets in the U.S. and found that the effect of market volatility is much stronger during bust periods. The stress that people experience during periods of poor market performance can be explained through uncertainty about their jobs, uncertainty about their future incomes, and uncertainty about their wealth, and thus impose strong negative effects on different aspects of life such as health outcomes and behavioral outcomes. Through the anticipatory theoretical framework, future market uncertainty should have an effect on individuals' contemporaneous life satisfaction. Therefore, this effect should have instantaneous effect and anticipatory effect on utility, measured by well-being. Our paper shows just that.

2. There is a possibility that market uncertainty, measured by the VIX, will only effect stockholders in the current period due to trading activity and therefore there is no anticipatory effect. In other words, through this explanation the effect of the VIX on well-being is only explained through traders whose well-beings are motivated by money made from trading activity. If this is the case, the effect of market uncertainty should not have an effect on their behaviors. Given this explanation, we propose another empirical framework to test whether the VIX affect different behaviors of the respondents. If we find that there is no effect on behavior, then any observed effects of the VIX on well-being are likely resulting from changes in income and wealth levels of market participants, not through changes in anticipatory utility. However, if there exist significant effects of the VIX on respondents' behaviors, we can confirm that the VIX does indeed affect well-being with the sample population, and not necessarily only the well-being of traders. In other words, if this is the case, we are convinced that our first explanation through the framework of anticipatory model is plausible. In addition, if we find significant effects of the VIX on well-being while controlling for stock ownership, we are confident that the VIX does capture the market stress and imposes the stress on individuals’ well-beings. 

3. There is also another possibility that the respondents are reporting their life satisfaction biasedly. They may be inclined to over report their well-being in the expectation that their well-being will improve. While they may wish to report a false sense of well-being for a myriad of reasons, previous research has shown that reported well-being and health in survey data are largely consistent with individuals' true states of beings. However, if this third explanation is true, or there are reporting errors coming from the survey, then we should not expect to find any significant effect of the VIX on any other health outcomes. This means that the VIX should not have an effect on behavioral factor, such as drinking. In other words, if we find that market uncertainty has no effect on behavioral factors then survey reporting errors are a plausible explanation for an observed significance in the relationship between the VIX and well-being. However, if there exists a significant effect of market uncertainty on these behavioral factors, then we are convinced that our anticipatory utility framework offers a plausible explanation.


# 3. Empirical framework & Descriptive Statistics

## 3.1 Empirical framework

Our baseline regression specification to examine the relation between stock market volatility and individuals’ well-being is:

$$ (1) Well-being_{i,s,t}=\beta_1MarketVolatility_t + \beta_2StockOwnership_{i,t} + \beta_XX_{i,s,t}+\tau_t+\gamma_s+\epsilon_{i,s,t}$$

$$ (2) ImpulsiveBehavior_{i,s,t}=\beta_1MarketVolatility_t + \beta_2StockOwnership_{i,t} + \beta_XX_{i,s,t}+\tau_t+\gamma_s+\epsilon_{i,s,t}$$


In model 1, the dependent variable is a measure of well-being. In this analysis, $Well-being_{i,s,t}$ is measured in reported General life satisfaction from the American general population.^[see Descriptive Statistics for detailed deconstruction for well-being]. These surveys span from 2013 to 2017 and are made by an individual $i$ in state $s$ on day $t$. $MarketVolatility_t$ is the percent change of the VIX, or the Implied Volatility Index, between two periods. The set of control variables, which may influence well-being, is suggested by previous work.^[see, e.g., Engelberg and Parsons, 2016; Cotti, Dunn, and Tefft, 2015; Davalos, Fang, and French, 2011; Fiuzat, Shaw, Thomas, Felker and O’Connor, 2010; Ruhm, 2005; Ruhm and Black, 2002] Particularly, $X_{i,s,t}$  is a matrix of individual-level demographic data including $age$, $gender$, $income$, $education-status$, $marital-status$, $employment-status$, $race$. In addition, $\tau_t$ are indicator variables for calendar months, and calendar years, $\gamma_s$ are state fixed effects. In model 2, $ImpulsiveBehaviors_{i,s,t}$ takes two a variable of interest, drinking, which indicates number of drinks in the last 30 days before the survey interview.

$\tau_t$ is a vector of month fixed effects and year fixed effects which accounts for the effect of seasonality on well-being behaviors. We also control for state fixed effects, $\gamma_s$, which would account for permanent differences across states that may affect well-being and its behaviors. These controls account for lifestyles patterns, and state infrastructures on life-satisfaction changes. In addition, adding these controls help account for unobserved factors that vary within states over time, such as social norms related to well-being behaviors. These also help us control for other state level time-varying factors such as changes in health care delivery services that closely follow tax revenues. The $\epsilon_{i,s,t}$ is an idiosyncratic random error term.

## 3.2 Descriptive Statistics

Our final data is presented in the most interpretable format. All variables and sub-categories are put into two different columns based on our main variable of interest, Well-being. The first column is 0, which presents respondents who reported "Dissatisfied" and "Very dissatisfied" while the second column, which is 1, presents respondents who reported "Satisfied", and "Very satisfied. This practice is done by previous works (Horn, Maclean, Strain 2016). While one may argue on the reliability of these subjective questions, previous works (Apouey and Clark 2015) suggests that they, in fact, capture not only an overall assessment of well-being but also combination of mental and physical health. In addition, Benjamins et al. 2004; Miilunpalo et al. 1997; Jylhä 2009 have shown that these measures can predict various health outcomes such as mortality and healthcare utilization. Thus, it is fair to conclude that the self-reported well-being measures from health surveys in our data plausibly correlate with objective health. Particularly, these self-reported measures from reliable sources such as the BRFSS have been used in many economics studies. ^[ see Maclean 2013; Apouey and Clark 2015; Maclean et al. 2015; Lenhart 2015; Bockerman and Ilmakunnas 2009; Gravelle and Sutton 2009; Lindeboom and van Doorslaer 2004; Cottini and Lucifora 2013] Although, some previous studies suggests there may exist some reporting errors (Baker, Stabile, and Deri 2004) which can affect our estimates, as described above, we try our best to minimize these errors by getting rid of incomplete respondents, only including variables that are in the main Panel, and getting rid of observations that are vaguely reported. ^[Depending on the questions, these are respondents that answer "Don't Know" or "Refused"]

In addition, our self-reported measures are particularly useful for a study of the short-run effects of market sentiments. It seems unlikely that more severe or objective measures of poor well-being conditions (e.g., mortality, chronic conditions, hospitalizations) will respond in the short run to a change in market volatility (and the associated income and time cost changes). Thus, our analysis of self-reported well-being, which captures how a person evaluates their well-being at a point in time, is potentially more responsive, and therefore more suitable, for our study objectives than more severe or objective measures. 

To provide a proper view of the dataset, we show the following chart, which represents the number of respondents that answered the survey throughout 2013-2017. Although there is a disproportion in number of participants who respond to the survey in the beginning of the year (January and February) compared to other months, our set of time control variables capture seasonality in our model analysis. Note that we do not show the descriptive statistics for the CPS data since they do not necessarily play a major role in our demographic variables. However, brief explanation of the CPS data is introduced at Section 4.


```{r, include=FALSE, echo=FALSE}
count.obs = BRFSS_FINANCE_data %>% count(yearmonth)
count.obs$yearmonth = as.yearmon(count.obs$yearmonth)
```

```{r,fig4, out.width = '70%', echo=FALSE, ,fig.align='center',fig.cap= 'Number of People Interviewed'}
# Number of respondents
plot(as.zoo(count.obs$yearmonth),as.zoo(count.obs$n),
     type='l',
     col = "steelblue",
     lwd = 2,
     ylab = "# Respondents",
     xlab = "Date",
     main = "Number of People Interviewed (Thousands)")
```


As discussed in the prior section, $MarketVolatility_t$ is the log returns of the VIX over the specified time frame, i.e. daily, weekly, bimonthly, and monthly returns. In performing this operation, we not only get an interpretable value (returns), but we also impose stationarity on each series. This allows us to derive the expected value of the VIX at time \emph{t}, $E[VIX_t]$. As shown in the graphs below, all $MarketVolatility_t$ series appear stationary, especially regarding the first moment. However, because market volatility tends to cluster into two distinct regimes, low volatility and high volatility, the S&P 500 returns exhibit non-constant variance over the full time series. We do not observe the same variance characteristics for the VIX, which appears to possess variance properties that are time-invariant, thus adhering more closely to the assumptions of our generalized linear estimation model. This fact provides additional empirical justification for considering the VIX as our primary variable of interest. Nevertheless, all return series, including all S&P 500 series, pass standard stationarity tests at the 5% level, as shown in the table below.

```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap= 'VIX Data'}
plot(dat_sentiment$date,as.zoo(dat_sentiment$VIX.rets),
     type = 'l',
     col = "steelblue",
     lwd = 2,
     ylab = "Log Returns",
     xlab = "Date",
     main = "Daily VIX Returns")

dat_sentiment$logVIXadjusted <- log(dat_sentiment$VIX.Adjusted)
plot(dat_sentiment$date,as.zoo(dat_sentiment$VIX.Adjusted),
     type = 'l',
     col = "steelblue",
     lwd = 2,
     ylab = "VIX Implied Annual Volatility",
     xlab = "Date",
     main = "VIX Daily Series")
plot(dat_sentiment$date,as.zoo(dat_sentiment$logVIXadjusted),
     type = 'l',
     col = "steelblue",
     lwd = 2,
     ylab = "VIX Implied Annual Volatility",
     xlab = "Date",
     main = "Log VIX Daily Series")
```


Shown below are Descriptive Statistics based on General Well-being. As presented, there is a larger number of respondents who report "Satisfied" and "Very satisfied" (N = 7,987) compared to those who report "Dissatisfied" and "Very dissatisfied" (N = 619). To put this in perspective, we show charts based on the variables for each group of our variable of interest.

\begingroup
\tabcolsep = 12.0pt
\def\arraystretch{0.65}
\captionsetup[table]{labelformat=empty,skip=1pt} 
{\footnotesize \begin{longtable}{llll}
\caption*{
\large Descriptive Statistics - Demographics\\ 
\small By General Health\\ 
} \\ 
\toprule
Characteristic & Statistic & 0, N = 530 & 1, N = 6,986 \\ 
\midrule
\textbf{Age} &  &  &  \\ 
\quad 18to24  & n / N (\%) & 3 / 530 (0.6\%) & 37 / 6986 (0.5\%) \\ 
\quad 25to34  & n / N (\%) & 13 / 530 (2.5\%) & 197 / 6986 (2.8\%) \\ 
\quad 35to44  & n / N (\%) & 39 / 530 (7.4\%) & 394 / 6986 (5.6\%) \\ 
\quad 45to54  & n / N (\%) & 122 / 530 (23\%) & 979 / 6986 (14\%) \\ 
\quad 55to64  & n / N (\%) & 199 / 530 (38\%) & 1837 / 6986 (26\%) \\ 
\quad 65older  & n / N (\%) & 154 / 530 (29\%) & 3542 / 6986 (51\%) \\ 
\textbf{Gender} &  &  &  \\ 
\quad female  & n / N (\%) & 329 / 530 (62\%) & 3932 / 6986 (56\%) \\ 
\quad male  & n / N (\%) & 201 / 530 (38\%) & 3054 / 6986 (44\%) \\ 
\textbf{Income} &  &  &  \\ 
\quad 50more  & n / N (\%) & 82 / 530 (15\%) & 2625 / 6986 (38\%) \\ 
\quad 15to25K  & n / N (\%) & 157 / 530 (30\%) & 1432 / 6986 (20\%) \\ 
\quad 25to35K  & n / N (\%) & 74 / 530 (14\%) & 915 / 6986 (13\%) \\ 
\quad 35to50K  & n / N (\%) & 40 / 530 (7.5\%) & 1061 / 6986 (15\%) \\ 
\quad le15K  & n / N (\%) & 177 / 530 (33\%) & 953 / 6986 (14\%) \\ 
\textbf{Education} &  &  &  \\ 
\quad COLgrad  & n / N (\%) & 116 / 530 (22\%) & 2107 / 6986 (30\%) \\ 
\quad attendCOL  & n / N (\%) & 176 / 530 (33\%) & 2094 / 6986 (30\%) \\ 
\quad HSgrad  & n / N (\%) & 160 / 530 (30\%) & 2143 / 6986 (31\%) \\ 
\quad K  & n / N (\%) & 78 / 530 (15\%) & 642 / 6986 (9.2\%) \\ 
\textbf{Marital Status} &  &  &  \\ 
\quad married  & n / N (\%) & 144 / 530 (27\%) & 3705 / 6986 (53\%) \\ 
\quad divorced  & n / N (\%) & 151 / 530 (28\%) & 1064 / 6986 (15\%) \\ 
\quad membermarriedcoup  & n / N (\%) & 16 / 530 (3.0\%) & 95 / 6986 (1.4\%) \\ 
\quad nevermarried  & n / N (\%) & 102 / 530 (19\%) & 737 / 6986 (11\%) \\ 
\quad separated  & n / N (\%) & 38 / 530 (7.2\%) & 132 / 6986 (1.9\%) \\ 
\quad widowed  & n / N (\%) & 79 / 530 (15\%) & 1253 / 6986 (18\%) \\ 
\textbf{Employment Status} &  &  &  \\ 
\quad wagesemployed  & n / N (\%) & 96 / 530 (18\%) & 2079 / 6986 (30\%) \\ 
\quad homemaker  & n / N (\%) & 14 / 530 (2.6\%) & 270 / 6986 (3.9\%) \\ 
\quad noworkless1  & n / N (\%) & 14 / 530 (2.6\%) & 119 / 6986 (1.7\%) \\ 
\quad noworkmore1  & n / N (\%) & 33 / 530 (6.2\%) & 129 / 6986 (1.8\%) \\ 
\quad retired  & n / N (\%) & 126 / 530 (24\%) & 3149 / 6986 (45\%) \\ 
\quad selfemployed  & n / N (\%) & 12 / 530 (2.3\%) & 411 / 6986 (5.9\%) \\ 
\quad student  & n / N (\%) & 2 / 530 (0.4\%) & 31 / 6986 (0.4\%) \\ 
\quad unable  & n / N (\%) & 233 / 530 (44\%) & 798 / 6986 (11\%) \\ 
\textbf{Race} &  &  &  \\ 
\quad white  & n / N (\%) & 429 / 530 (81\%) & 5932 / 6986 (85\%) \\ 
\quad asian  & n / N (\%) & 1 / 530 (0.2\%) & 43 / 6986 (0.6\%) \\ 
\quad black  & n / N (\%) & 68 / 530 (13\%) & 763 / 6986 (11\%) \\ 
\quad native  & n / N (\%) & 25 / 530 (4.7\%) & 142 / 6986 (2.0\%) \\ 
\quad other  & n / N (\%) & 7 / 530 (1.3\%) & 99 / 6986 (1.4\%) \\ 
\quad pacific  & n / N (\%) & 0 / 530 (0\%) & 7 / 6986 (0.1\%) \\ 
\hline 
\multicolumn{4}{l}{\textit{Note: This table does not include State Descriptive Statistics}}\\
\bottomrule
\end{longtable}}
\endgroup



\begingroup
\tabcolsep = 12.0pt
\def\arraystretch{0.65}
\captionsetup[table]{labelformat=empty,skip=1pt}
\begin{longtable}{lccccccc}
\caption*{
\large Descriptive Statistics - Chronic Conditions\\ 
} \\ 
\toprule
Statistic & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{St. Dev.} & \multicolumn{1}{c}{Min} & \multicolumn{1}{c}{Pctl(25)} & \multicolumn{1}{c}{Pctl(75)} & \multicolumn{1}{c}{Max} \\ 
\hline 
\midrule
Health Insured & 7,516 & 0.949 & 0.219 & 0 & 1 & 1 & 1 \\ 
Heart Disease & 7,516 & 0.142 & 0.349 & 0 & 0 & 0 & 1 \\ 
Arthritis & 7,516 & 0.493 & 0.500 & 0 & 0 & 1 & 1 \\ 
Stroke & 7,516 & 0.085 & 0.280 & 0 & 0 & 0 & 1 \\ 
Asthma & 7,516 & 0.155 & 0.362 & 0 & 0 & 0 & 1 \\ 
Bronchitis & 7,516 & 0.125 & 0.331 & 0 & 0 & 0 & 1 \\ 
Depression & 7,516 & 0.266 & 0.442 & 0 & 0 & 1 & 1 \\ 
Cancer & 7,516 & 0.143 & 0.350 & 0 & 0 & 0 & 1 \\ 
Diabetes & 7,516 & 0.915 & 0.280 & 0 & 1 & 1 & 1 \\ 
\hline 
\bottomrule
\end{longtable}
\endgroup


# 4. Methodology and Empirical Results
## 4.1 Methodology

One challenging part of our analysis is to achieve stock-holding data from the Current Population Survey. Unlike the BRFSS, where all of the variables are categorical, variables from the CPS are continuous. Thus, we have to recode all of the demographics variables from the CPS to match with the BRFSS because we want to maintain the satisfaction variable in the BRFSS while merging all the demographics from both the CPS and the BRFSS together. The CPS data has a crucial element for our analysis, in which it consists of the stock ownership variable. In specific, having stock-holding variable helps us to define the relationship of market uncertainty on the stock-holding population and non-stock-holding population in our sample. 

Our next plan of action is to achieve the propensity to own stock from the Current Population Survey (CPS), then merge this information onto the BRFSS survey, together with the Financial data. We employ logistic regression to investigate the relationship of some demographic variables on whether or not the respondents are holding stock. Suggested by previous literature (Kreinin et al 1959), we use some of the most significant demographic predictors of stock ownership, including $age$, $gender$, $income$, $education-status$, and $race$, in our regression to achieve the propensity of stock ownership at individual levels and then merge this propensity score onto the BRFSS dataset based on the demographics mentioned above. ^[See Appendix for the Logistic regression result of demographics on stock ownership.]

Our propensity result indicates that males are more likely to own stocks compared to females. Compared to respondents who have a college degree, respondents who has not graduate from college, only have high school diploma, and has not graduated from high school are much less likely to own stocks. Respondents who make at least 50 thousand dollars a year are the most likely to own stock. Respondents in the age group of 18-24 are the least likely to own stocks while respondents who are 65 and older are the most likely to own stocks. Causation population are the most likely to own stocks, follows by mixed race and Asian. In general sense, our result is a decent representation of the truth and thus indicates that our model for Stock Ownership is pretty accurate. 


```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap= 'CPS Demographics - Age and Race' }
cpsasec_dfs.bothsampling$ownstock.graph <- ifelse(cpsasec_dfs.bothsampling$ownstock==1,"Stock", "No Stock")
#Bar-plot
age.plot.cps <- ggplot(cpsasec_dfs.bothsampling, aes(x = factor(ownstock.graph), fill = factor(calculated_age)))
age.plot.cps + geom_bar(position="dodge") + labs(x = "Stock Ownership", y = "Count") + scale_fill_brewer(palette="Accent",name = "Age Brackets") 

race.plot.cps <- ggplot(cpsasec_dfs.bothsampling, aes(x = factor(ownstock.graph), fill = factor(calculated_race)))
race.plot.cps + geom_bar(position="dodge") + labs(x = "Stock Ownership", y = "Count") + scale_fill_brewer(palette="Accent",name = "Race") 

```

```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap=  'CPS Demographics - Education and Income'}
education.plot.cps <- ggplot(cpsasec_dfs.bothsampling, aes(x = factor(ownstock.graph), fill = factor(calculated_education)))
education.plot.cps + geom_bar(position="dodge") + labs(x = "Stock Ownership", y = "Count") + scale_fill_brewer(palette="Accent",name = "Education") 

income.plot.cps <- ggplot(cpsasec_dfs.bothsampling, aes(x = factor(ownstock.graph), fill = factor(calculated_income)))
income.plot.cps + geom_bar(position="dodge") + labs(x = "Stock Ownership", y = "Count") + scale_fill_brewer(palette="Accent",name = "Income") 
```

```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap=  'CPS Demographics - Gender and Employment Status'}
gender.plot.cps <- ggplot(cpsasec_dfs.bothsampling, aes(x = factor(ownstock.graph), fill = factor(gender)))
gender.plot.cps + geom_bar(position="dodge") + labs(x = "Stock Ownership", y = "Count") + scale_fill_brewer(palette="Accent",name = "Gender") 
```
`

```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap=  'Propensity Score for Stock Ownership'}

hist(cpsasec_dfs.bothsampling$psvalue.cps,main="Histogram for Propensity of Stock Ownership", 
     xlab="Propensity Score", 
     border="black", 
     col="grey")

#latex.cps=stargazer(glm.ps.cps, type="latex", title="Regression Results for Stock Ownership")

```

\begingroup
\tabcolsep = 12.0pt
\def\arraystretch{0.65}
\captionsetup[justification=centering]{labelformat=empty,skip=1pt} 
{\scriptsize \begin{longtable}[c]{lc} 
\caption*{Regression Results for Stock Ownership\\ 
\small \textit{Dependent variable: General Well-being} \\ 
} \\ 
\toprule \centering
& Stock Ownership\\ 
\\[-1.8ex] & (1) \\
\midrule 
\hline \\[-1.8ex] 
 gendermale & 0.242$^{***}$ \\ 
  & (0.011) \\ 
 
 calculated\_educationattendCOL & $-$0.954$^{***}$ \\ 
  & (0.014) \\ 
 
 calculated\_educationHSgrad & $-$1.411$^{***}$ \\ 
  & (0.014) \\ 
 
 calculated\_educationK & $-$2.371$^{***}$ \\ 
  & (0.028) \\ 
 
 calculated\_income15to25K & $-$1.244$^{***}$ \\ 
  & (0.027) \\ 
 
 calculated\_income25to35K & $-$0.972$^{***}$ \\ 
  & (0.024) \\ 
 
 calculated\_income35to50K & $-$0.724$^{***}$ \\ 
  & (0.019) \\ 
 
 calculated\_incomele15K & $-$1.568$^{***}$ \\ 
  & (0.030) \\ 
 
 calculated\_age25to34 & 0.661$^{***}$ \\ 
  & (0.029) \\ 
 
 calculated\_age35to44 & 1.152$^{***}$ \\ 
  & (0.028) \\ 
 
 calculated\_age45to54 & 1.471$^{***}$ \\ 
  & (0.028) \\ 
 
 calculated\_age55to64 & 1.752$^{***}$ \\ 
  & (0.028) \\ 
 
 calculated\_age65older & 2.183$^{***}$ \\ 
  & (0.028) \\ 
 
 calculated\_raceasian & $-$0.287$^{***}$ \\ 
  & (0.022) \\ 
 
 calculated\_raceblack & $-$0.933$^{***}$ \\ 
  & (0.021) \\ 
 
 calculated\_racenative & $-$0.405$^{***}$ \\ 
  & (0.057) \\ 
 
 calculated\_raceother & $-$0.153$^{***}$ \\ 
  & (0.043) \\ 
 
 calculated\_racepacific & $-$0.717$^{***}$ \\ 
  & (0.088) \\ 
 
 Constant & $-$0.385$^{***}$ \\ 
  & (0.028) \\ 
 
\hline \\[-1.8ex] 
Observations & 177,296 \\ 
Log Likelihood & $-$98,710.070 \\ 
Akaike Inf. Crit. & 197,458.100 \\ 
\hline 
\hline \\[-1.8ex] 
\multicolumn{2}{l}{Note: Panel consists of the 2013-2017 survey sample waves of CPS.}\\
\multicolumn{2}{l}{Model estimates with a Logistic Regression. Model controls for}\\
\multicolumn{2}{l}{demographics, including age, gender, race, income, education status,}\\
\multicolumn{2}{l}{marital status,employment status.  See Appendix for detailed variable}\\
\multicolumn{2}{l}{definition, data source, and construction.}\\
\multicolumn{2}{l}{ $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\bottomrule
\end{longtable}}
\endgroup

The analysis is proceeded with the generalized linear probability model (logistics). All the regressions are weighted using the BRFSS sampling weights. As described above, controlling for a set of demographics variables is extremely useful in generating precise estimates, given different demographics in our survey. In addition, rather than using adjusted closing price data for the VIX as many previous works, we want to set our main variables of interest in a stationary form so that not only we do not have to treat for unit-root problems within out independent variables but also we do not have to worry about the fact that the upward performance trends of the S&P500 would bias our results. In addition, by controlling for State Fixed-Effects, Monthly Fixed-Effects, and Yearly Fixed-Effects, we aim to get to the most precise estimates possible. Respectively, these additional controls account for the impact of seasonality that may exist in some of the behaviors such as physical activity (Ruhm 2005), permanent differences across states that may affect health and health behaviors, such as lifestyles patterns, state infrastructures on health care, and confounding factors that may trend linearly. As mentioned above, adding these helps account for unobserved factors that vary within states over time, such as social norms related to health behaviors. They also help us control for other state level time-varying factors such as changes in health care delivery services that closely follow tax revenues.

We face quite a difficult task when there is a huge imbalance in classification for our main dependent variable, general well-being. After re-coding the data class 0 for respondents who report "Dissatisfied" and "Very Dissatisfied" and class 1 for respondents who report "Satisfied" and "Very Satisfied", our dependent variable was categorized with much less observations of class 0 and much more observations of 1. In such cases, when the binary response variable is imbalanced in the data so the proportion of class 1 heavily dominates the class 0 proportion, the logistic model does not perform well for detection of the outcome with low proportion. In other words, the reporting Specificity of the model would be very low, which means the probability of detection of class 1 cases would be unsatisfyingly low. Thus, they result in biased predictions and misleading accuracy measures. there are several methods available to tackle imbalanced classification problems, however they are not easy to perform in a large dataset such as ours. To combat the issue in this analysis, we impose two different methods, then compare these results to see how robust our models are. The first method is the combination of both under-sampling and oversampling where we under-sample class 1 and over-sample class 0 to produce a perfect proportion of 50/50 for class 1 and class 0. The problem with this first model is that the data generated from oversampling have a certain amount of repeated observations and the data generated from under-sampling is deprived of important information in the original data. This leads to possible inaccuracies in the resulting performance. The second method is to simply impose under-sampling for class 1, in which we would have to randomly cut the sample of class 1 to get to the same amount of observations as class 0. The problem with this second method is that we would lose a significant information from the sample of class 1, although it keeps all the information in class 0. This method leaves just enough observations for us to analyze the data. However, by imposing the two methods and by properly comparing the two results, we should yield estimates in the middle range of biases, which produce arguably the best and most accurate estimates. Both of these techniques are commonly used in previous literature using imbalanced classification logistic models, as they would improve the prediction accuracy significantly. Each sampling method has its own advantages and disadvantages. In this paper, we perform a robust analysis using both sampling methods. 

Panel below presents the results of the VIX on reported well-being. We estimate logistic regressions to establish our first \textbf{Hypothesis} - increases in the VIX, i.e. when the market is under stress or uncertainty, is negatively associated with reported well-being. Running on three different data samples for different sampling methods, we find evidence that there are significant negative effects linking market volatility to well-being. Specifically, running logistic regressions on general well-being outcome predicting 1, or being satisfied, we find significant negative effects on all three different regressions, signaling the probability of the VIX increasing and having better well-being will decrease compared to bad well-being. 

In model 2, we perform a different econometric approach for the main outcome, the likelihood of drinking. We refer this variable as alcohol consumption for the rest of the analysis. After re-coding the data class 0 for respondents who report to have no drink and class 1 for respondents who report having some drinks, our dependent binary variable was slightly unbalanced. As explained above, in such cases when the binary response variable is imbalanced in the sample data so that the proportion of class 1 heavily dominates the class 0 proportion, then logistic models do not perform well for detection of the outcome with low proportion. In other words, the reporting Specificity of the model would be very low, which means the probability of detection of class 1 cases would be unsatisfyingly low. Thus, to tackle imbalanced classification problems, we impose different sampling methods to achieve our estimates. See Appendix for after-sampling descriptive statistics.

In addition, by nature of the question in the BRFSS on our alcohol consumption variable, which reports "Adults who reported having had at least one drink of alcohol in the past 30 days," we want to lag our VIX variable by 30 days. This process is done on our Financial data to ensure that we do not mix up the date on our final merged data. As mentioned above, we seek to perform a similar analysis from Sias 2017 in regard to the volatility. Our independent variable of interest, instead of taking in the return series of the VIX, uses the natural log of the monthly lag of the VIX adjusted price. 


## 4.2 Empirical Results

### Model 1: Estimating the relationship between the VIX and Life Satisfaction

Analysis for Model 1 will impose a combination of oversampling and under sampling method, which leaves 7,516 observations. We can also perform oversampling for our results as well. However, the downside of oversampling is that the data will be significantly inflated and may potentially create biases. Performing under sampling is not the best practice as it leaves too little observations to perform statistical analysis. Therefore, a combination of oversampling and under sampling is our best approach as we maintain the perfect amount of observations from class 1, and slightly inflate class 0. 

Visually, we have fantastic controls, which include gender, age with 6 different categories, income with 5 different categories, education status with 4 categories, marital status with 6 different categories, race with 6 different categories, chronic health conditions, and more importantly employment status with 8 different categories. Moreover, a set of fixed effects for months, years, and states ensures that we can capture the effect with least errors as possible. Although this study does not necessarily focus on the effect of demographics on life satisfaction nor the effect of chronic conditions on life satisfaction, our results show that these effects across the board are expected.

Across 6 different models, the first three models (1)-(3) takes the daily return series of the VIX while the last three models (3)-(6) takes the weekly return models. Model (1) and model (4) do not control for chronic conditions while models (2) and (5) does. Model (3) and (6) control for $psvalue.cps$, which is the propensity score value for stock ownership. Our results indicate that the effects of the VIX on well-being are pretty consistent overall. The magnitudes of the results do not fluctuate significantly across different specifications. Model (3) and model (6) are our prime models, where they control for chronic conditions, propensity score of stock ownership, and sample weights. The interaction term shows that as market is under stress, increases in the likelihood of the respondents owning stock result to a decrease in the probability of respondents feeling satisfied. This makes sense because this directly shows the income effect in the regression. In other words, as the market is performing poorly, the more likely the respondents hold stocks, the more likely they have poor life satisfaction. In addition, we want to ensure to include our weekly models to show that the results are consistent with respondents behaviors. It is possible that the daily return series of the VIX do not necessarily affect respondents if they do not even have time to check on their positions of their stock holding on the interview day, especially for stockholders, but by setting the VIX series to weekly, we are confident that this series should capture the effect of market volatility on well-being. 

\begingroup
\tabcolsep = 12.0pt
\def\arraystretch{0.65}
\captionsetup[justification=centering]{labelformat=empty,skip=1pt} 
{\scriptsize \begin{longtable}[c]{lccc} 
\caption*{Regression Results for Well-Being and Daily VIX - Under and Over Sampling Combination\\ 
\small \textit{Dependent variable: General Well-being} \\ 
} \\ 
\toprule \centering
& Daily Rets Model & Daily Rets Model & Daily Rets Model\\ 
  &  No Chronic &  With Chronic & Interaction\\ 
\\[-1.8ex] & (1) & (2) & (3)\\
\midrule 
\hline \\[-1.8ex] 
 VIX.rets & $-$1.376$^{***}$ & $-$0.803$^{***}$ & $-$0.448$^{***}$ \\ 
  & (0.055) & (0.060) & (0.118) \\ 
  
 psvalue.cps &  &  & $-$1.672$^{***}$ \\ 
  &  &  & (0.109) \\ 

 VIX.rets:psvalue.cps &  &  & $-$0.858$^{***}$ \\ 
  &  &  & (0.262) \\ 
  
 calculated\_age25to34 & $-$0.786$^{***}$ & $-$0.604$^{***}$ & $-$0.417$^{***}$ \\ 
  & (0.047) & (0.050) & (0.053) \\ 
  
 calculated\_age35to44 & $-$1.224$^{***}$ & $-$1.054$^{***}$ & $-$0.799$^{***}$ \\ 
  & (0.045) & (0.047) & (0.051) \\ 
  
 calculated\_age45to54 & $-$0.824$^{***}$ & $-$0.631$^{***}$ & $-$0.302$^{***}$ \\ 
  & (0.044) & (0.046) & (0.052) \\ 
  
 calculated\_age55to64 & $-$0.881$^{***}$ & $-$0.541$^{***}$ & $-$0.118$^{**}$ \\ 
  & (0.044) & (0.046) & (0.054) \\ 
  
 calculated\_age65older & $-$0.310$^{***}$ & $-$0.134$^{***}$ & 0.423$^{***}$ \\ 
  & (0.044) & (0.047) & (0.060) \\ 
  
 gendermale & 0.006 & $-$0.261$^{***}$ & $-$0.173$^{***}$ \\ 
  & (0.008) & (0.009) & (0.011) \\ 
  
 calculated\_income15to25K & $-$1.127$^{***}$ & $-$0.979$^{***}$ & $-$1.410$^{***}$ \\ 
  & (0.012) & (0.013) & (0.031) \\ 
  
 calculated\_income25to35K & $-$1.354$^{***}$ & $-$1.265$^{***}$ & $-$1.618$^{***}$ \\ 
  & (0.013) & (0.014) & (0.027) \\ 
  
 calculated\_income35to50K & $-$0.096$^{***}$ & $-$0.020 & $-$0.280$^{***}$ \\ 
  & (0.013) & (0.015) & (0.022) \\ 
  
 calculated\_incomele15K & $-$1.071$^{***}$ & $-$0.748$^{***}$ & $-$1.262$^{***}$ \\ 
  & (0.014) & (0.015) & (0.037) \\ 
  
 calculated\_educationattendCOL & $-$0.202$^{***}$ & $-$0.186$^{***}$ & $-$0.529$^{***}$ \\ 
  & (0.010) & (0.011) & (0.025) \\ 
  
 calculated\_educationHSgrad & $-$0.177$^{***}$ & $-$0.087$^{***}$ & $-$0.573$^{***}$ \\ 
  & (0.010) & (0.011) & (0.034) \\ 
  
 calculated\_educationK & $-$0.098$^{***}$ & 0.014 & $-$0.660$^{***}$ \\ 
  & (0.015) & (0.016) & (0.047) \\ 
  
 marital\_statusdivorced & $-$0.869$^{***}$ & $-$0.930$^{***}$ & $-$0.938$^{***}$ \\ 
  & (0.010) & (0.011) & (0.011) \\ 
  
 marital\_statusmembermarriedcoup & $-$1.070$^{***}$ & $-$1.560$^{***}$ & $-$1.553$^{***}$ \\ 
  & (0.027) & (0.029) & (0.029) \\ 
  
 marital\_statusnevermarried & $-$0.653$^{***}$ & $-$0.925$^{***}$ & $-$0.922$^{***}$ \\ 
  & (0.013) & (0.014) & (0.014) \\ 
  
 marital\_statusseparated & $-$1.512$^{***}$ & $-$1.921$^{***}$ & $-$1.921$^{***}$ \\ 
  & (0.020) & (0.021) & (0.021) \\ 
  
 marital\_statuswidowed & $-$0.395$^{***}$ & $-$0.870$^{***}$ & $-$0.882$^{***}$ \\ 
  & (0.013) & (0.014) & (0.014) \\ 
  
 employment\_statushomemaker & 0.714$^{***}$ & 0.764$^{***}$ & 0.783$^{***}$ \\ 
  & (0.021) & (0.024) & (0.024) \\ 
  
 employment\_statusnoworkless1 & $-$0.951$^{***}$ & $-$0.527$^{***}$ & $-$0.546$^{***}$ \\ 
  & (0.023) & (0.024) & (0.024) \\ 
  
 employment\_statusnoworkmore1 & $-$1.784$^{***}$ & $-$1.385$^{***}$ & $-$1.357$^{***}$ \\ 
  & (0.021) & (0.023) & (0.023) \\ 
  
 employment\_statusretired & $-$0.152$^{***}$ & $-$0.011 & $-$0.007 \\ 
  & (0.012) & (0.013) & (0.013) \\ 
  
 employment\_statusselfemployed & 0.554$^{***}$ & 0.660$^{***}$ & 0.669$^{***}$ \\ 
  & (0.020) & (0.022) & (0.022) \\ 
  
 employment\_statusstudent & $-$0.076 & 0.219$^{***}$ & 0.131$^{**}$ \\ 
  & (0.053) & (0.057) & (0.058) \\ 
  
 employment\_statusunable & $-$1.111$^{***}$ & $-$0.510$^{***}$ & $-$0.513$^{***}$ \\ 
  & (0.011) & (0.013) & (0.013) \\ 
  
 calculated\_raceasian & 1.660$^{***}$ & 0.651$^{***}$ & 0.554$^{***}$ \\ 
  & (0.081) & (0.078) & (0.079) \\ 
  
 calculated\_raceblack & 0.441$^{***}$ & 0.509$^{***}$ & 0.277$^{***}$ \\ 
  & (0.011) & (0.012) & (0.020) \\ 
  
 calculated\_racenative & $-$1.217$^{***}$ & $-$1.087$^{***}$ & $-$1.157$^{***}$ \\ 
  & (0.028) & (0.029) & (0.030) \\ 
  
 calculated\_raceother & 1.162$^{***}$ & 0.977$^{***}$ & 0.911$^{***}$ \\ 
  & (0.033) & (0.037) & (0.038) \\ 
  
 calculated\_racepacific & 12.140 & 11.213 & 11.338 \\ 
  & (26.207) & (17.602) & (19.943) \\ 
  
 health\_insured &  & 0.580$^{***}$ & 0.579$^{***}$ \\ 
  &  & (0.015) & (0.015) \\ 
  
 heart\_disease &  & $-$0.235$^{***}$ & $-$0.235$^{***}$ \\ 
  &  & (0.012) & (0.012) \\ 
  
 arthritis &  & $-$0.314$^{***}$ & $-$0.306$^{***}$ \\ 
  &  & (0.009) & (0.009) \\ 
  
 stroke &  & $-$0.522$^{***}$ & $-$0.550$^{***}$ \\ 
  &  & (0.013) & (0.013) \\ 
  
 asthma &  & $-$0.112$^{***}$ & $-$0.109$^{***}$ \\ 
  &  & (0.011) & (0.011) \\ 
  
 bronchitis &  & $-$0.163$^{***}$ & $-$0.170$^{***}$ \\ 
  &  & (0.012) & (0.012) \\ 
  
 depression &  & $-$1.588$^{***}$ & $-$1.589$^{***}$ \\ 
  &  & (0.009) & (0.009) \\ 
  
 cancer &  & 0.189$^{***}$ & 0.188$^{***}$ \\ 
  &  & (0.012) & (0.012) \\ 
  
 diabetes &  & $-$0.455$^{***}$ & $-$0.454$^{***}$ \\ 
  &  & (0.017) & (0.017) \\ 
  
 Constant & 2.704$^{***}$ & 3.203$^{***}$ & 4.073$^{***}$ \\ 
  & (0.050) & (0.058) & (0.082) \\ 
  
\hline \\[-1.8ex] 
Month FE  & Yes & Yes & Yes  \\ 
State FE  & Yes & Yes & Yes \\ 
Year FE  & Yes & Yes & Yes \\ 
\hline \\[-1.8ex] 
Survey Weights & Yes & Yes & Yes \\ 
Observations  & 7,516 & 7,516 & 7,516 \\ 
\hline 
\hline \\[-1.8ex] 
\multicolumn{4}{l}{Note: Panel consists of the 2013-2017 survey sample waves of BRFSS. BRFSS sample weights applied. All models}\\
\multicolumn{4}{l}{estimated with a Logistic Regression. All models control for demographics, including age, gender, race, income,}\\
\multicolumn{4}{l}{education status, marital status, employment status.  See Appendix for detailed variable definition, data source,}\\
\multicolumn{4}{l}{and construction.}\\
\multicolumn{4}{l}{ $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\bottomrule
\end{longtable}}
\endgroup

Panel shows the effects of the VIX on well-being outcomes. Holding all else constant, an additional increase in the VIX return is associated with an increase in the odds ratio of reporting better well-being, relative to poor well-being. Respectively, these effects persist with the same negative effects and similar magnitudes across different sampling methods. This means that as the market is in under-stress or in a state of uncertainty, it is more likely for the respondents from the sample to fall into the categories representing dissatisfaction, compared to the categories representing satisfaction. Given the standard errors of the VIX, we are confident to reject the null hypothesis, concluding there is a negative association between the VIX and reported well-being. In addition, the 95% confidence interval lies within the negative zone of the effect, which also confirms that this effect is negatively correlated. 

To better understand the effects, we want to investigate the partial effects of the VIX on well-being based on the minimum, mean, median, and the maximum of the propensity of stockholding. By simply takes the derivative of specification (3) with respect to the VIX, we should be able to achieve the partial effects of the VIX, dependent on the values of stockholding propensity. Our results show that at the minimum of the propensity score of stockholding, VIX still imposes a negative effect on well-being. This means that even when the respondents do not own stock, market stress still negatively impact the well-beings of the respondents. As the propensity to own stock increases, this effect also increases in the magnitude.


\begin{table}[!htbp] \centering 
\caption{Daily Partial Effects of VIX on Well-being}
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{$\frac{\delta Well-being}{\delta VIX.rets}$}     & \multicolumn{4}{c|}{\textbf{Propensity Score of Stock-holding}}       \\ \cline{2-5} 
                                      & \textit{Minimum} & \textit{Median} & \textit{Mean} & \textit{Maximum} \\ \hline
\textit{\textbf{Daily VIX Partial Effects}} & $-0.4568488^{***}$       & $-0.7606199^{***}$      & $-0.7938655^{***}$    & $-1.207772^{***}$        \\ \hline
\end{tabular}
\end{table}

The daily partial effects of the VIX return series at the minimum, median, mean, and maximum propensity of owning stock are at -0.46, -0.76, -0.79, and -1.2 respectively. At the mean of stockholding propensity, additional increase in the VIX daily return is associated with a decrease in well-being by an expected log count of 0.76, holding all else constant, and it is significant at the 1\% level. The effect does change given different propensity of holding stock. Again, the magnitude of this negative effect gets larger if the propensity of owning stock is larger.

As mentioned above, there is a possibility that by the time the respondents have the interview dates, the stockholders may not be aware of the daily VIX series and thus do not get affected immediately. By including our weekly return series of the VIX, we should be able to capture the full effect of the market turbulence on individual well-being. We purposefully do not include the monthly returns for two main reasons. First, fluctuations in the market during any certain month are unexpected. The VIX monthly return series can fluctuate up and down all the time and by just capturing the return difference between the first day of the month and the last day of the month, we are excluding all the fluctuations in between, where we are sure that these fluctuations within any certain month are the most important information that we want to use. Therefore, we do not want to use the monthly return series. Second, given that our final dataset, after a combination of up and down sampling, only consists of roughly 7,000 observations, by using monthly data, we would end up with roughly 1,000 observations, which are not enough to perform our statistical analysis with extremely dense controls.


\begingroup
\tabcolsep = 12.0pt
\def\arraystretch{0.65}
\captionsetup[justification=centering]{labelformat=empty,skip=1pt} 
{\scriptsize \begin{longtable}[c]{lccc} 
\caption*{Regression Results for Well-Being and Weekly VIX - Under and Over Sampling Combination\\ 
\small \textit{Dependent variable: General Well-being} \\ 
} \\ 
\toprule \centering
& Weekly Rets Model & Weekly Rets Model & Weekly Rets Model\\ 
  &  No Chronic &  With Chronic & Interaction\\ 
\\[-1.8ex] & (4) & (5) & (6)\\
\midrule 
\hline \\[-1.8ex] 
 VIX.weeklyrets & $-$1.946$^{***}$ & $-$1.927$^{***}$ & $-$0.375$^{**}$ \\ 
  & (0.085) & (0.105) & (0.179) \\ 
  
 psvalue.cps &  &  & $-$3.431$^{***}$ \\ 
  &  &  & (0.501) \\ 

 VIX.weeklyrets:psvalue.cps &  &  & $-$3.771$^{***}$ \\ 
  &  &  & (0.359) \\ 
  
 calculated\_age25to34 & $-$16.108 & $-$13.545 & $-$12.704 \\ 
  & (64.152) & (85.391) & (65.212) \\ 
  
 calculated\_age35to44 & $-$17.786 & $-$14.710 & $-$13.562 \\ 
  & (64.152) & (85.391) & (65.212) \\ 
  
 calculated\_age45to54 & $-$16.468 & $-$13.529 & $-$12.278 \\ 
  & (64.152) & (85.391) & (65.212) \\ 
  
 calculated\_age55to64 & $-$16.539 & $-$13.168 & $-$11.743 \\ 
  & (64.152) & (85.391) & (65.212) \\ 
  
 calculated\_age65older & $-$14.947 & $-$12.112 & $-$10.426 \\ 
  & (64.152) & (85.391) & (65.212) \\ 
  
 gendermale & 0.188$^{***}$ & 0.106$^{***}$ & 0.298$^{***}$ \\ 
  & (0.026) & (0.034) & (0.041) \\ 
  
 calculated\_income15to25K & $-$0.629$^{***}$ & $-$1.541$^{***}$ & $-$2.469$^{***}$ \\ 
  & (0.037) & (0.048) & (0.140) \\ 
  
 calculated\_income25to35K & $-$0.285$^{***}$ & $-$0.909$^{***}$ & $-$1.675$^{***}$ \\ 
  & (0.041) & (0.051) & (0.121) \\ 
  
 calculated\_income35to50K & $-$0.320$^{***}$ & $-$0.766$^{***}$ & $-$1.376$^{***}$ \\ 
  & (0.039) & (0.050) & (0.097) \\ 
  
 calculated\_incomele15K & $-$1.406$^{***}$ & $-$2.206$^{***}$ & $-$3.304$^{***}$ \\ 
  & (0.043) & (0.059) & (0.171) \\ 
  
 calculated\_educationattendCOL & $-$0.018 & 0.124$^{***}$ & $-$0.551$^{***}$ \\ 
  & (0.031) & (0.038) & (0.109) \\ 
  
 calculated\_educationHSgrad & 0.319$^{***}$ & 0.658$^{***}$ & $-$0.335$^{**}$ \\ 
  & (0.031) & (0.041) & (0.153) \\ 
  
 calculated\_educationK & 0.604$^{***}$ & 1.511$^{***}$ & 0.096 \\ 
  & (0.045) & (0.058) & (0.221) \\ 
  
 marital\_statusdivorced & $-$1.365$^{***}$ & $-$0.978$^{***}$ & $-$0.975$^{***}$ \\ 
  & (0.030) & (0.040) & (0.040) \\ 
  
 marital\_statusmembermarriedcoup & $-$0.899$^{***}$ & $-$1.790$^{***}$ & $-$1.772$^{***}$ \\ 
  & (0.079) & (0.099) & (0.099) \\ 
  
 marital\_statusnevermarried & $-$1.024$^{***}$ & $-$1.144$^{***}$ & $-$1.147$^{***}$ \\ 
  & (0.040) & (0.051) & (0.051) \\ 
  
 marital\_statusseparated & $-$1.967$^{***}$ & $-$1.188$^{***}$ & $-$1.088$^{***}$ \\ 
  & (0.056) & (0.063) & (0.067) \\ 
  
 marital\_statuswidowed & 0.488$^{***}$ & 0.014 & $-$0.006 \\ 
  & (0.040) & (0.052) & (0.052) \\ 
  
 employment\_statushomemaker & 0.490$^{***}$ & 0.979$^{***}$ & 1.030$^{***}$ \\ 
  & (0.074) & (0.088) & (0.088) \\ 
  
 employment\_statusnoworkless1 & $-$1.877$^{***}$ & $-$0.100 & $-$0.068 \\ 
  & (0.081) & (0.098) & (0.099) \\ 
  
 employment\_statusnoworkmore1 & $-$2.118$^{***}$ & $-$0.842$^{***}$ & $-$0.812$^{***}$ \\ 
  & (0.061) & (0.080) & (0.080) \\ 
  
 employment\_statusretired & $-$1.374$^{***}$ & 0.168$^{***}$ & 0.205$^{***}$ \\ 
  & (0.040) & (0.054) & (0.055) \\ 
  
 employment\_statusselfemployed & 1.744$^{***}$ & 2.726$^{***}$ & 2.796$^{***}$ \\ 
  & (0.089) & (0.111) & (0.111) \\ 
  
 employment\_statusunable & $-$1.707$^{***}$ & $-$0.088$^{*}$ & $-$0.086$^{*}$ \\ 
  & (0.038) & (0.050) & (0.050) \\ 
  
 calculated\_raceasian & 10.580 & 12.119 & 10.029 \\ 
  & (154.651) & (420.384) & (154.651) \\ 
  
 calculated\_raceblack & 0.453$^{***}$ & 0.757$^{***}$ & 0.268$^{***}$ \\ 
  & (0.039) & (0.055) & (0.088) \\ 
  
 calculated\_racenative & 0.546$^{***}$ & 1.587$^{***}$ & 1.450$^{***}$ \\ 
  & (0.073) & (0.079) & (0.083) \\ 
  
 calculated\_raceother & 2.584$^{***}$ & 2.323$^{***}$ & 2.092$^{***}$ \\ 
  & (0.099) & (0.119) & (0.124) \\ 
  
 calculated\_racepacific & $-$1.312 & 9.058 & 6.650 \\ 
  & (221.569) & (582.780) & (221.879) \\ 
  
 health\_insured &  & 0.775$^{***}$ & 0.782$^{***}$ \\ 
  &  & (0.069) & (0.070) \\ 
  
 heart\_disease &  & $-$0.605$^{***}$ & $-$0.607$^{***}$ \\ 
  &  & (0.040) & (0.040) \\ 
  
 arthritis &  & $-$0.787$^{***}$ & $-$0.766$^{***}$ \\ 
  &  & (0.033) & (0.033) \\ 
  
 stroke &  & $-$0.813$^{***}$ & $-$0.829$^{***}$ \\ 
  &  & (0.043) & (0.043) \\ 
  
 asthma &  & 0.728$^{***}$ & 0.701$^{***}$ \\ 
  &  & (0.041) & (0.041) \\ 
  
 bronchitis &  & $-$1.118$^{***}$ & $-$1.194$^{***}$ \\ 
  &  & (0.039) & (0.039) \\ 
  
 depression &  & $-$2.954$^{***}$ & $-$2.956$^{***}$ \\ 
  &  & (0.033) & (0.033) \\ 
  
 cancer &  & $-$0.321$^{***}$ & $-$0.368$^{***}$ \\ 
  &  & (0.036) & (0.037) \\ 
  
 diabetes &  & $-$1.429$^{***}$ & $-$1.436$^{***}$ \\ 
  &  & (0.081) & (0.082) \\ 
  
 Constant & 20.081 & 21.973 & 23.079 \\ 
  & (64.152) & (85.391) & (65.213) \\ 
    
\hline \\[-1.8ex] 
Month FE  & Yes & Yes & Yes  \\ 
State FE  & Yes & Yes & Yes \\ 
Year FE  & Yes & Yes & Yes \\ 
\hline \\[-1.8ex] 
Survey Weights & Yes & Yes & Yes \\ 
Observations & 1,397 & 1,397 & 1,397 \\ 
\hline 
\hline \\[-1.8ex] 
\multicolumn{4}{l}{Note: Panel consists of the 2013-2017 survey sample waves of BRFSS. BRFSS sample weights applied. All models}\\
\multicolumn{4}{l}{estimated with a Logistic Regression. All models control for demographics, including age, gender, race, income,}\\
\multicolumn{4}{l}{education status, marital status, employment status.  See Appendix for detailed variable definition, data source,}\\
\multicolumn{4}{l}{and construction.}\\
\multicolumn{4}{l}{ $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\bottomrule
\end{longtable}}
\endgroup

\begin{table}[!htbp] \centering 
\caption{Weekly Partial Effects of VIX on Well-being}
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{$\frac{\delta Well-being}{\delta VIX.rets}$}     & \multicolumn{4}{c|}{\textbf{Propensity Score of Stock-holding}}       \\ \cline{2-5} 
                                      & \textit{Minimum} & \textit{Median} & \textit{Mean} & \textit{Maximum} \\ \hline
\textit{\textbf{VIX Weekly Partial Effects}} & $-0.3890383^{***}$       & $-0.8971497^{***}$      & $-0.9527591^{***}$    & $-1.645091^{***}$        \\ \hline
\end{tabular}
\end{table}

Although the weekly model does not give us as much observations as the daily model, we still find our results to be quite consistent. Our Weekly Partial Effect table indicates that the results does not fluctuate as much compared to the daily partial effects of the VIX on well-being. This weekly model can act as our first robustness check as it not only includes different specifications but also uses the weekly series rather than the daily series, which they are arguably difficult to fully capture the daily effect of the VIX on well-being on the interview date. 

### Model 2: Estimating the relationship between the VIX and Impulsive Behaviors.

This section dedicates to our another hypothesis, which we hypothesize that increase in the VIX would results to more likely to consume alcohol consumption. Our regression models are based on previous literature of the VIX on smoking and drinking (Kalcheva, McLemore, Sias 2017). Regression takes the dependent variable of drinking, which is a  binary outcomes, and the main independent variable of interest, the natural log of the VIX adjusted price, which is a stationary series. We also take an additional control for the likelihood of stock ownership.



\begingroup
\tabcolsep = 12.0pt
\def\arraystretch{0.65}
\captionsetup[justification=centering]{labelformat=empty,skip=1pt} 
{\scriptsize \begin{longtable}[c]{lccc} 
\caption*{Regression Results for Alcohol Consumption and Monthly VIX - Under and Over Sampling Combination\\ 
\small \textit{Dependent variable: General Well-being} \\ 
} \\ 
\toprule \centering
& Monthly Log VIX & Monthly Log VIX & Monthly Log VIX\\ 
  &  No Chronic &  With Chronic & Interaction\\ 
\\[-1.8ex] & (1) & (2) & (3)\\
\midrule 
\hline \\[-1.8ex] 
 log(monthlag.VIX.Adjusted) & 0.240$^{***}$ & 0.253$^{***}$ & 0.187$^{***}$ \\ 
  & (0.024) & (0.024) & (0.042) \\ 
  
 psvalue.cps &  &  & $-$1.904$^{***}$ \\ 
  &  &  & (0.214) \\ 

 log(monthlag.VIX.Adjusted):psvalue.cps &  &  & 0.173$^{**}$ \\ 
  &  &  & (0.070) \\ 
  
 calculated\_age25to34 & $-$0.574$^{***}$ & $-$0.610$^{***}$ & $-$0.494$^{***}$ \\ 
  & (0.051) & (0.050) & (0.052) \\ 
  
 calculated\_age35to44 & $-$0.452$^{***}$ & $-$0.363$^{***}$ & $-$0.152$^{***}$ \\ 
  & (0.048) & (0.048) & (0.051) \\ 
  
 calculated\_age45to54 & $-$0.737$^{***}$ & $-$0.553$^{***}$ & $-$0.265$^{***}$ \\ 
  & (0.048) & (0.048) & (0.053) \\ 
  
 calculated\_age55to64 & $-$0.808$^{***}$ & $-$0.574$^{***}$ & $-$0.206$^{***}$ \\ 
  & (0.048) & (0.048) & (0.055) \\ 
  
 calculated\_age65older & $-$1.185$^{***}$ & $-$0.880$^{***}$ & $-$0.392$^{***}$ \\ 
  & (0.049) & (0.049) & (0.060) \\ 
  
 gendermale & 0.453$^{***}$ & 0.528$^{***}$ & 0.595$^{***}$ \\ 
  & (0.008) & (0.008) & (0.009) \\ 
  
 calculated\_income15to25K & $-$0.988$^{***}$ & $-$0.983$^{***}$ & $-$1.363$^{***}$ \\ 
  & (0.012) & (0.012) & (0.030) \\ 
  
 calculated\_income25to35K & $-$0.749$^{***}$ & $-$0.744$^{***}$ & $-$1.047$^{***}$ \\ 
  & (0.012) & (0.012) & (0.025) \\ 
  
 calculated\_income35to50K & $-$0.561$^{***}$ & $-$0.563$^{***}$ & $-$0.787$^{***}$ \\ 
  & (0.011) & (0.011) & (0.019) \\ 
  
 calculated\_incomele15K & $-$1.182$^{***}$ & $-$1.190$^{***}$ & $-$1.631$^{***}$ \\ 
  & (0.014) & (0.015) & (0.035) \\ 
  
 calculated\_educationattendCOL & $-$0.140$^{***}$ & $-$0.151$^{***}$ & $-$0.441$^{***}$ \\ 
  & (0.009) & (0.009) & (0.023) \\ 
  
 calculated\_educationHSgrad & $-$0.481$^{***}$ & $-$0.466$^{***}$ & $-$0.887$^{***}$ \\ 
  & (0.010) & (0.010) & (0.032) \\ 
  
 calculated\_educationK & $-$0.608$^{***}$ & $-$0.604$^{***}$ & $-$1.207$^{***}$ \\ 
  & (0.016) & (0.016) & (0.046) \\ 
  
 marital\_statusdivorced & 0.315$^{***}$ & 0.335$^{***}$ & 0.323$^{***}$ \\ 
  & (0.011) & (0.011) & (0.011) \\ 
  
 marital\_statusmembermarriedcoup & 0.480$^{***}$ & 0.485$^{***}$ & 0.494$^{***}$ \\ 
  & (0.033) & (0.033) & (0.033) \\ 
  
 marital\_statusnevermarried & $-$0.052$^{***}$ & $-$0.052$^{***}$ & $-$0.040$^{***}$ \\ 
  & (0.013) & (0.013) & (0.013) \\ 
  
 marital\_statusseparated & 0.382$^{***}$ & 0.435$^{***}$ & 0.448$^{***}$ \\ 
  & (0.023) & (0.023) & (0.023) \\ 
  
 marital\_statuswidowed & 0.247$^{***}$ & 0.272$^{***}$ & 0.268$^{***}$ \\ 
  & (0.012) & (0.012) & (0.012) \\ 
  
 employment\_statushomemaker & $-$0.189$^{***}$ & $-$0.223$^{***}$ & $-$0.196$^{***}$ \\ 
  & (0.019) & (0.020) & (0.020) \\ 
  
 employment\_statusnoworkless1 & $-$0.041 & $-$0.019 & $-$0.016 \\ 
  & (0.029) & (0.029) & (0.029) \\ 
  
 employment\_statusnoworkmore1 & $-$0.229$^{***}$ & $-$0.180$^{***}$ & $-$0.152$^{***}$ \\ 
  & (0.027) & (0.027) & (0.027) \\ 
  
 employment\_statusretired & 0.052$^{***}$ & 0.078$^{***}$ & 0.083$^{***}$ \\ 
  & (0.011) & (0.011) & (0.011) \\ 
  
 employment\_statusselfemployed & $-$0.028$^{*}$ & $-$0.049$^{***}$ & $-$0.044$^{***}$ \\ 
  & (0.015) & (0.015) & (0.015) \\ 
  
 employment\_statusstudent & $-$0.520$^{***}$ & $-$0.430$^{***}$ & $-$0.487$^{***}$ \\ 
  & (0.056) & (0.056) & (0.056) \\ 
  
 employment\_statusunable & $-$0.816$^{***}$ & $-$0.669$^{***}$ & $-$0.664$^{***}$ \\ 
  & (0.013) & (0.014) & (0.014) \\ 
  
 calculated\_raceasian & $-$1.286$^{***}$ & $-$1.235$^{***}$ & $-$1.343$^{***}$ \\ 
  & (0.040) & (0.040) & (0.041) \\ 
  
 calculated\_raceblack & 0.094$^{***}$ & 0.057$^{***}$ & $-$0.168$^{***}$ \\ 
  & (0.012) & (0.012) & (0.020) \\ 
  
 calculated\_racenative & $-$0.172$^{***}$ & $-$0.100$^{***}$ & $-$0.193$^{***}$ \\ 
  & (0.025) & (0.025) & (0.026) \\ 
  
 calculated\_raceother & $-$0.930$^{***}$ & $-$0.981$^{***}$ & $-$1.012$^{***}$ \\ 
  & (0.031) & (0.031) & (0.031) \\ 
  
 calculated\_racepacific & 3.331$^{***}$ & 3.452$^{***}$ & 3.362$^{***}$ \\ 
  & (0.353) & (0.353) & (0.354) \\ 
  
 health\_insured &  & $-$0.273$^{***}$ & $-$0.269$^{***}$ \\ 
  &  & (0.016) & (0.016) \\ 
  
 heart\_disease &  & $-$0.263$^{***}$ & $-$0.264$^{***}$ \\ 
  &  & (0.011) & (0.011) \\ 
  
 arthritis &  & 0.011 & 0.013$^{*}$ \\ 
  &  & (0.008) & (0.008) \\ 
  
 stroke &  & $-$0.405$^{***}$ & $-$0.407$^{***}$ \\ 
  &  & (0.014) & (0.014) \\ 
  
 asthma &  & 0.033$^{***}$ & 0.038$^{***}$ \\ 
  &  & (0.010) & (0.010) \\ 
  
 bronchitis &  & $-$0.180$^{***}$ & $-$0.186$^{***}$ \\ 
  &  & (0.012) & (0.012) \\ 
  
 depression &  & 0.033$^{***}$ & 0.030$^{***}$ \\ 
  &  & (0.009) & (0.009) \\ 
  
 cancer &  & $-$0.094$^{***}$ & $-$0.097$^{***}$ \\ 
  &  & (0.011) & (0.011) \\ 
  
 diabetes &  & $-$0.412$^{***}$ & $-$0.419$^{***}$ \\ 
  &  & (0.014) & (0.014) \\ 
  
 Constant & 0.697$^{***}$ & 1.072$^{***}$ & 2.027$^{***}$ \\ 
  & (0.085) & (0.087) & (0.135) \\ 
  
\hline \\[-1.8ex] 
Month FE  & Yes & Yes & Yes  \\ 
State FE  & Yes & Yes & Yes \\ 
Year FE  & Yes & Yes & Yes \\ 
\hline \\[-1.8ex] 
Survey Weights & Yes & Yes & Yes \\ 
Observations & 7,376 & 7,376 & 7,376 \\ 
\hline 
\hline \\[-1.8ex] 
\multicolumn{4}{l}{Note: Panel consists of the 2013-2017 survey sample waves of BRFSS. BRFSS sample weights applied. All models}\\
\multicolumn{4}{l}{estimated with a Logistic Regression. All models control for demographics, including age, gender, race, income,}\\
\multicolumn{4}{l}{education status, marital status, employment status.  See Appendix for detailed variable definition, data source,}\\
\multicolumn{4}{l}{and construction.}\\
\multicolumn{4}{l}{ $^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\bottomrule
\end{longtable}}
\endgroup


\begin{table}[!htbp] \centering 
\caption{Partial Monthly Effects of VIX on Alcohol Consumption}
\begin{tabular}{|c|c|c|c|c|}
\hline
\multirow{2}{*}{$\frac{\delta Well-being}{\delta VIX.rets}$}     & \multicolumn{4}{c|}{\textbf{Propensity Score of Stock-holding}}       \\ \cline{2-5} 
                                      & \textit{Minimum} & \textit{Median} & \textit{Mean} & \textit{Maximum} \\ \hline
\textit{\textbf{Log VIX Partial Effects}} & $0.1882297^{***}$       & $0.2689059^{***}$      & $0.2692511^{***}$    & $0.3396178^{***}$        \\ \hline
\end{tabular}
\end{table}


These results indicate the rejection of the null hypothesis and conclude that there are positive association between the VIX and impulsive behaviors. Specifically, when the market is under-stress, or as the VIX adjusted price increases by one percent, alcohol consumption increases by an expected log count of 0.24 and 0.253 in models (1) and (2). All of these estimates are significant at the 1% level, holding all else constant. The partial effects of the VIX on alcohol consumption in specification (3) follows the same pattern in our Model 1 of the VIX and well-being. To fully figure out the true effect of model (3) in the above specification, we want to take the derivative of specification (3) with respect to the natural log of the VIX. We find that the partial effects at minimum, median, mean, and maximum propensity of stock ownership are approximately 0.19, 0.27, 0.27, and 0.34 respectively. As the propensity score of owning stocks increases, the magnitude of the effect also increases. The estimates fit very nicely into our main model of anticipatory framework in which we have shown that future market uncertainty does indeed have effects on contemporaneous behaviors. In other words, we provide evidence against the notion that the VIX only affects the well-being individuals who benefit from stockholding but does not affect their behaviors.



## 4.3 Goodness of Fit and Additional Robustness Check

```{r, include=FALSE, echo=FALSE}
#model 
test.glm.vix.both = glm(well_being ~  VIX.rets*psvalue.cps + calculated_age + gender + calculated_income + calculated_education + marital_status + employment_status + calculated_race +state_name.x + imonth + iyear.x + health_insured + heart_disease + arthritis + stroke + asthma + bronchitis + depression + cancer + diabetes, weights=xstrwt, family=binomial(link=logit),    data = trainingData)
sum.test.glm.vix.both=summary(test.glm.vix.both)
sum.test.glm.vix.both

predicted <- predict(test.glm.vix.both, testData, type="response")
optCutOff <- optimalCutoff(testData$well_being, predicted)[1]
misClassError(testData$well_being, predicted, threshold = optCutOff)
plotROC(testData$well_being, predicted)  #0.82, which is pretty good
sensitivity(testData$well_being, predicted, threshold = optCutOff) 
specificity(testData$well_being, predicted, threshold = optCutOff) 
confusionMatrix(testData$well_being, predicted, threshold = optCutOff)
```

Based on previous literature, we originally did not control for chronic conditions. However, we demonstrate the robustness of our results by controlling for a series of additional controls, including the propensity of stock ownership, and especially chronic conditions. Our results show that our models are very robust and consistent across different models of the VIX return series, as well as the VIX adjusted price series. During the analysis, we suspect that fluctuation in the return series in the implied volatility, VIX, could come from the actual performance of the stock market overall and thus we wanted to control for the return series of stock performance. Some of the best indicators of stock performance that are widely used are the S&P 500 (SPY), the Dow Jones Industrial Average (DJI), and the Russell 2000 (RUT). However, there are two main reasons for why we do not include them in our model. First, the VIX is actually derived from the S&P 500 and therefore, they are highly correlated. Second, Welch two sample t-test between the SPY returns and VIX returns gives the p-value of 0, indicating that the true difference in mean of the two series is equal to 0. Similarly, we run the same t-test between the VIX returns series against the DJI return series and the RUT return series and find that the p-values are also at 0. This means that the difference in mean between the VIX return series and the DJI and RUT series are significantly equal to 0. We also run the correlation test between these variables and find that the VIX and SPY returns are 81\% negatively correlated, the VIX and RUT returns are 71\% negatively correlated, and the VIX and DJI returns are 77\% negatively correlated. Therefore, we did not include these variables in our regressions.  

In addition, to assess the accuracy of the predictions generated by our models, we calculated a variety of goodness of fit test, including Sensitivity, Specificity, Positive Predictive Value, Negative Predictive Value, and most importantly the Concordance of the models. Specificity is the percentage of 0’s (actuals) correctly predicted while Sensitivity (or True Positive Rate) is the percentage of 1’s (actuals) correctly predicted by the model. Specificity can also be calculated as (1 - False Positive Rate). Ideally, the model-calculated-probability-scores of all actual Positive’s, (Ones) should be greater than the model-calculated-probability-scores of ALL the Negatives (Zeroes). Such a model is said to be perfectly concordant and a highly reliable one, in which this phenomenon can be measured by Concordance and Discordance. In simpler words, of all combinations of 1-0 pairs (actuals), Concordance is the percentage of pairs whose scores of actual positives are greater than the scores of actual negatives. For a perfect model, this will be 100%. Although yielding a 100% concordance is nearly impossible for any model, the higher the concordance the better. 

In addition, we also plot the ROC Curve (Receiver Operating Characteristics Curve) to show the ability of the model to classify the difference between class 0 and class 1. Receiver Operating Characteristics Curve traces the percentage of true positives accurately predicted by a given logit model as the prediction probability cutoff is lowered from 1 to 0. For a good model, as the cutoff is lowered, it should mark more of actual 1’s as positives and lesser of actual 0’s as 1’s. Therefore, for a good model, the curve should rise fast, indicating that the TPR (Y-Axis) increases faster than the FPR (X-Axis) as the cutoff score decreases. The greater the area under the ROC curve, better the predictive ability of the model. In simpler terms, the closer the ROC value is to 1, or 100%, the better the model. 


```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap= 'ROC Curves VIX Model'}
plotROC(testData$well_being, predicted) 
```


\begin{table}[!htbp] \centering
\caption{Confusion Matrix - VIX Daily Returns} 
  \label{}
\begin{tabular}{|c|c|c|c|c|}
\hline
Confusion Matrix                    & \multicolumn{3}{c|}{\textbf{Actuals}}                  & \multirow{2}{*}{Training N = 5,208} \\ \cline{1-4}
\multirow{3}{*}{\textbf{Predicted}} & Classification & 0                 & 1                 &                            \\ \cline{2-5} 
                                    & 0              & 896             & 335              & NPV = 73\%                 \\ \cline{2-5} 
                                    & 1              & 221              & 856             & PPV = 79\%                 \\ \hline
\multicolumn{2}{|c|}{Testing N = 2,308}                      & Specificity: 80\% & Sensitivity: 72\% & \textbf{Concordance: 83\%}        \\ \hline
\end{tabular}
\end{table}

The plot above represents the ROC curve and the Confusion Matrix for VIX model using daily returns. As explained above, our model is fitted with a Concordance of 82%, indicating a quality model. In addition, with Specificity and Sensitivity ranging from 72% to 80% from the Confusion Matrix, we are confident that the performance of the model on the test data is exceptional. Generally, there are trade-offs between sensitivity and specificity, in which re-sampling the data to improve sensitivity can significantly reduce the specificity level. Therefore, having the standard estimates of the Specificity and Sensitivity as we do shows that our model is not only robust, but also accurate (as shown in the ROC curves).

# 5. Conclusion

This study explores the impact of stock market returns on self-reported well-being. Our results show clear patterns, similar to previous research, in which self-reported well-being is worsened during periods of market turmoil and uncertainty. Using market volatility, or the VIX, we find evidence that self-reported well-being is more likely to be reported in the category of “Satisfied” and “Very Satisfied” compared to the category of “Dissatisfied” and “Very Dissatisfied”  when the implied volatility index declines, or when the market volatility indicates a relative decline in uncertainty. This study is novel in that we employ expected volatility as our primary variable of interest, as opposed to other mainstream stock market indicators, such as the S&P 500 or the Dow Jones Industrial Average, and thus we assess the relationship between forward-looking volatility expectations and individual well-being metrics. Furthermore, since the future economic conditions are pertinent for both non-stockholders or stockholders, it is expected that human responses are widespread and not simply restricted to individuals actively participating in the market. In sum, with various tests and robustness checks, our results strongly support our hypothesis on the VIX and confirm previous evidence on mainstream market indicators, such as the Dow Jones. In addition, we are able to fit our main model into the anticipatory framework, and show that the VIX return series, acting as an anticipatory index, influence life satisfaction of survey respondents, which acts as our main utility measure. This is also a novel approach that takes the association of the financial market with well-being and life satisfaction into a new level, where research has spent too much time investigating the relationship of human behaviors on the stock market but they have completely forgotten about the opposite effects. We hope that this study can add to the behavioral economic and modern finance literature in that particular point of view.

Previous research has shown that the effect mechanism between financial markets and well-being is derived from several possible factors. Earlier research (Brenner & Mooney, 1983; Catalano & Dooley, 1983) suggests that the level of stress due to market conditions may lead to self-medication. Risky health behaviors can also be the result of market downturns. Behaviors such as smoking, overeating, and binge drinking are more likely to occur when market performance is poorer (Colman & Dave, 2011; Cotti & Tefft, 2011; 23 Ruhm & Black, 2002; Ruhm, 2005). In addition, Cotti, Dunn, Tefft 2013 found a diminished income effect when assessing the impact of the Dow Jones on health, suggesting that market and economic stress plays a role in one’s inclination to participate in risky health behaviors. Therefore, our findings help explain why behavioral biases are more severe when expected market volatility is high (Kumar, 2009). 

While the study estimates are strongly investigated, there are several limitations to this study. First, although we show a deep channel of how market volatility affects well-being, we cannot conclude that this effect is causal. Income effect indeed plays a big role into the effect of market volatility, especially when this effect gets larger when the propensity of holding stock is higher. Nonetheless, we are not able to fully fix for potential endogeneity issues. In fact, we take our result of how market volatility influences well-being in a indirect way. Individuals can be affected by market uncertainty in many ways, market crashes, potential job loss, etc. Second, although we show that the VIX can act as an anticipatory index for the market uncertainty, we do not fully understand the link of the index, which is very artificial and can be arguably easily manipulated, that connects to human behaviors. The stock market in the recent years are not like the stock market a couple of decades ago, where it was pretty selective in who participates. Nowadays, teenagers with a social security number can participate in the stock market and the behaviors coming from these inexperienced players can inflate the market significantly. Third, we understand that we impose many different sampling methods to tackle class imbalance for our well-being and life satisfaction variable. As a result, we do not believe that our estimates can perfectly predict the relationship of market uncertainty and life satisfaction. This is quite intuitive given that we had to use a total of three different sampling tricks to get our efficient and consistent estimates. 

Nonetheless, the paper underscores the exciting cross-section fields of behavioral economic, finance and health. Although the study is inspired by a hand-full of previous literature, we are unaware of research that are similar or the same as our study. By understanding the implication of stock market on behavior, and life satisfaction, this paper not only helps shed lights into scientific studies but also helps motivate policy makers to navigate and impose necessary restrictions on the financial stock market, which can have serious consequences to human well-being and life satisfaction.




\newpage
### Appendix

In this analysis, we refer well-being variable to $Well-being$. For demographic variables, given the nature of the BRFSS survey, all of them are categorical variables. $gender$ is recoded from "sex" in BRFSS from 2013-2017. The question asked "What is your sex? or What was your sex at birth? Was it…" We recoded this variable with 1 being "Male", 0 being "Female" and got rid of "Don’t know/Not Sure", and "Refused". Variable $marital\_status$ is recoded from section "Demographics", under label "Marital Status." This variable is a categorical variable with "Married", "Divorced", "Widowed", "Separated", "Never married", and "A member of an unmarried couple". We got rid of options "Refused", and missing data for this variable. Variable $employment\_status$ is also under "Demographics" section. The question asked "Question: Are you currently…?" and the answers ranging from "Employed for wages", "Self-employed", "Out of work for 1 year or more", "Out of work for less than 1 year", "A homemaker", "A student", "Retired", "Unable to work." Options "Refused" and “BLANK” (missing data) are eliminated from the analysis. Variable $calculated\_race$ is a calculated race variable with 6 categories "White", "Black or African American", "American Indian or Alaskan Native", "Asian", "Native Hawaiian or other Pacific Islander" and "Other race." I got rid of "No preferred race", "Don’t know/Not sure", "Refused" and missing data. Variable $calculated\_education$ represents education status with 4 main categories "Did not graduate High School", "Graduated High School", "Attended College or Technical School" and "Graduated from College or Technical School." We also got rid of "Don’t know/Not sure/Missing" value. Variable $calculated\_income$ represents income brackets from "Less than $\$15,000$", "$\$15,000$ to less than $\$25,000$", "$\$25,000$ to less than $\$35,000$", "$\$35,000$ to less than $\$50,000$", and "$\$50,000$ or more." Category "Don’t know/Not sure/Missing" is eliminated. Variable $calculated\_age$ represent calculated age variable with categories "Age 18 to 24", "Age 25 to 34", "Age 35 to 44", "Age 45 to 54", "Age 55 to 64", "Age 65 or older."  Missing data is also eliminated. 

The CPS survey spans from 2013 to 2017. Almost all variables used are continuous and thus we have to recode them to be categorical. Stock ownership, $ownstock$, is recoded from the question which ask “A anytime during 20.., did you have shares of stock in corporations or mutual funds?” We recoded this variable with 1 being “Yes”, 0 being “No”, and got rid of “Not in universe.” $gender$ is recoded with 1 being “Male”, 0 being “Female.” $calculated\_education$ is recoded from a question regarding “education attainment.” In specific, “Did not graduate High School” which takes in “Less than 1st grade”, “1st, 2nd, 3rd, or 4th grade”, “5th or 6th grade”, “7th and 8th grade”, “9th grade”, “10th grade”, “11th grade” and “12th grade no diploma” in the CPS. “Graduated High School” takes “High school graduate - high school diploma or equivalent” in the CPS. “Attended College or Technical School” takes “Some college but no degree”, “Associate degree in college - occupation/vocation program”, and “Associate degree in college - academic program” in the CPS. “Graduated from College or Technical School” takes “Bachelor's degree (for example: BA, AB, BS)”, “Master's degree (for example: MA, MS, MENG, MED, MSW, MBA)”, “Professional school degree (for example: MD, DDS, DVM, LLB, JD)” and “Doctorate degree (for example: PHD, EDD).” Variable $calculated\_age$ is recoded from a continuous age variable from the CPS to fit in the age categories of the BRFSS, which are “Age 18 to 24”, “Age 25 to 34”, “Age 35 to 44”, “Age 45 to 54”, “Age 55 to 64”, “Age 65 or older.” Variable $calculated\_income$ is also recoded from a continuous income variable from the CPS to fit in the income categories in the BRFSS, which are “Less than \$15,000”, “\$15,000 to less than \$25,000”, “\$25,000 to less than \$35,000”, “\$35,000 to less than \$50,000”, and “\$50,000 or more.” Variable $calculated\_race$ is a calculated race variable with 6 categories “White”, “Black or African American”, “American Indian or Alaskan Native”, “Asian”, “Native Hawaiian or other Pacific Islander” and “Other race.”


Down below are the descriptive statistics for our final dataset after sampling. Sampling method includes a combination of over and under sampling. By imposing this method, we are confident that our results are consistent and not so far off from results from solely over sampling or under sampling.

```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap= 'After Sampling: Demographics - Age and Race' }
aftersampling.descriptives <- data_balanced_both
aftersampling.descriptives$well_being.graph <- ifelse(aftersampling.descriptives$well_being==1,"Satisfied", "Not Satisfied")
#Bar-plot
age.plot <- ggplot(aftersampling.descriptives, aes(x = factor(well_being.graph), fill = factor(calculated_age)))
age.plot + geom_bar(position="dodge") + labs(x = "Self-Reported General Well-being", y = "Count") + scale_fill_brewer(palette="Accent",name = "Age Brackets") 

race.plot <- ggplot(aftersampling.descriptives, aes(x = factor(well_being.graph), fill = factor(calculated_race)))
race.plot + geom_bar(position="dodge") + labs(x = "Self-Reported General Well-being", y = "Count") + scale_fill_brewer(palette="Accent",name = "Race") 


```

```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap=  'After Sampling: Demographics - Education and Income'}
education.plot <- ggplot(aftersampling.descriptives, aes(x = factor(well_being.graph), fill = factor(calculated_education)))
education.plot + geom_bar(position="dodge") + labs(x = "Self-Reported General Well-being", y = "Count") + scale_fill_brewer(palette="Accent",name = "Education") 

income.plot <- ggplot(aftersampling.descriptives, aes(x = factor(well_being.graph), fill = factor(calculated_income)))
income.plot + geom_bar(position="dodge") + labs(x = "Self-Reported General Well-being", y = "Count") + scale_fill_brewer(palette="Accent",name = "Income") 
```

```{r, out.width = '.49\\linewidth',fig.show='hold',fig.align='center', echo=FALSE, fig.cap=  'After Sampling: Demographics - Gender and Employment Status'}
gender.plot <- ggplot(aftersampling.descriptives, aes(x = factor(well_being.graph), fill = factor(gender)))
gender.plot + geom_bar(position="dodge") + labs(x = "Self-Reported General Well-being", y = "Count") + scale_fill_brewer(palette="Accent",name = "Gender") 

employment.plot <- ggplot(aftersampling.descriptives, aes(x = factor(well_being.graph), fill = factor(employment_status)))
employment.plot + geom_bar(position="dodge") + labs(x = "Self-Reported General Well-being", y = "Count") + scale_fill_brewer(palette="Accent",name = "Employment Status") 
```


\newpage
### Reference

Apouey, Benedicte, and Andrew E. Clark. “Winning Big but Feeling No Better? The Effect of Lottery Prizes on Physical and Mental Health.” Health Economics 24, no. 5 (2015): 516–38. https://doi.org/10.1002/hec.3035.

Baker, Michael, Mark Stabile, and Catherine Deri. “What Do Self-Reported, Objective, Measures of Health Measure?” Journal of Human Resources XXXIX, no. 4 (October 2, 2004): 1067–93. https://doi.org/10.3368/jhr.XXXIX.4.1067.

Benjamin, Emelia J., Martin G. Larson, Michelle J. Keyes, Gary F. Mitchell, Ramachandran S. Vasan, John F. Keaney, Birgitta T. Lehman, Shuxia Fan, Ewa Osypiuk, and Joseph A. Vita. “Clinical Correlates and Heritability of Flow-Mediated Dilation in the Community: The Framingham Heart Study.” Circulation 109, no. 5 (February 10, 2004): 613–19. https://doi.org/10.1161/01.CIR.0000112565.60887.1E.

Böckerman, Petri, and Pekka Ilmakunnas. “Unemployment and Self-Assessed Health: Evidence from Panel Data.” Health Economics 18, no. 2 (2009): 161–79. https://doi.org/10.1002/hec.1361.

Brenner, M. Harvey, and Anne Mooney. “Unemployment and Health in the Context of Economic Change.” Social Science & Medicine 17, no. 16 (1983): 1125–38. https://doi.org/10.1016/0277-9536(83)90005-9.

Carleton, R. Nicholas. “The Intolerance of Uncertainty Construct in the Context of Anxiety Disorders: Theoretical and Practical Perspectives.” Expert Review of Neurotherapeutics 12, no. 8 (August 1, 2012): 937–47. https://doi.org/10.1586/ern.12.82.

Catalano, Ralph A., and David Dooley. “Health Effects of Economic Instability: A Test of Economic Stress Hypothesis.” Journal of Health and Social Behavior 24, no. 1 (1983): 46–60. https://doi.org/10.2307/2136302.

Cotti, Chad, Richard A. Dunn, and Nathan Tefft. “The Dow Is Killing Me: Risky Health Behaviors and the Stock Market.” Health Economics 24, no. 7 (2015): 803–21. https://doi.org/10.1002/hec.3062.

Cotti, Chad, and Nathan Tefft. “Decomposing the Relationship between Macroeconomic Conditions and Fatal Car Crashes during the Great Recession: Alcohol- and Non-Alcohol-Related Accidents in: The B.E. Journal of Economic Analysis & Policy Volume 11 Issue 1 (2011).” Accessed March 30, 2020. https://www.degruyter.com/view/journals/bejeap/11/1/article-bejeap.2011.11.1.2860.xml.xml.

Cottini, Elena, and Claudio Lucifora. “Mental Health and Working Conditions in Europe.” ILR Review 66, no. 4 (July 1, 2013): 958–88. https://doi.org/10.1177/001979391306600409.

Dávalos, María E., Hai Fang, and Michael T. French. “Easing the Pain of an Economic Downturn: Macroeconomic Conditions and Excessive Alcohol Consumption.” Health Economics 21, no. 11 (November 2012): 1318–35. https://doi.org/10.1002/hec.1788.

Dave, Dhaval M, Jennifer Tennant, and Gregory J Colman. “Isolating the Effect of Major Depression on Obesity: Role of Selection Bias.” Working Paper. Working Paper Series. National Bureau of Economic Research, May 2011. https://doi.org/10.3386/w17068.

Deaton, Angus S. “The Financial Crisis and the Well-Being of Americans.” Working Paper. Working Paper Series. National Bureau of Economic Research, June 2011. https://doi.org/10.3386/w17128.

Engelberg, Joseph, and Christopher A. Parsons. “Worrying about the Stock Market: Evidence from Hospital Admissions.” The Journal of Finance 71, no. 3 (2016): 1227–50. https://doi.org/10.1111/jofi.12386.

Fiuzat, Mona, Linda K. Shaw, Laine Thomas, G. Michael Felker, and Christopher M. O’Connor. “United States Stock Market Performance and Acute Myocardial Infarction Rates in 2008-2009 (from the Duke Databank for Cardiovascular Disease).” The American Journal of Cardiology 106, no. 11 (December 1, 2010): 1545–49. https://doi.org/10.1016/j.amjcard.2010.07.027.

Goidel, Kirby, Stephen Procopio, Dek Terrell, and H. Denis Wu. “Sources of Economic News and Economic Expectations.” American Politics Research 38, no. 4 (July 1, 2010): 759–77. https://doi.org/10.1177/1532673X09355671.

Gravelle, Hugh, and Matt Sutton. “Income, Relative Income, and Self-Reported Health in Britain 1979–2000.” Health Economics 18, no. 2 (2009): 125–45. https://doi.org/10.1002/hec.1354.

Horn, Brady P., Johanna Catherine Maclean, and Michael R. Strain. “Do Minimum Wage Increases Influence Worker Health?” Economic Inquiry 55, no. 4 (2017): 1986–2007. https://doi.org/10.1111/ecin.12453.

Jylhä, Marja. “What Is Self-Rated Health and Why Does It Predict Mortality? Towards a Unified Conceptual Model.” Social Science & Medicine (1982) 69, no. 3 (August 2009): 307–16. https://doi.org/10.1016/j.socscimed.2009.05.013.

Kahneman, Daniel, and Amos Tversky. “Prospect Theory: An Analysis of Decision under Risk.” Econometrica 47, no. 2 (1979): 263–91. https://doi.org/10.2307/1914185.

Kumar, Alok. “Who Gambles in the Stock Market? - KUMAR - 2009 - The Journal of Finance - Wiley Online Library,” 2009. https://onlinelibrary.wiley.com/doi/abs/10.1111/j.1540-6261.2009.01483.x.

Lenhart, Amanda, Aaron Smith, Monica Anderson, Maeve Duggan, and Andrew Perrin. “Teens, Technology and Friendships,” 2015. https://www.semanticscholar.org/paper/Teens%2C-technology-and-friendships-Lenhart-Smith/f34b35eef23822a5466c7f75270ad37b59072c2b.

Lindeboom, Maarten, and Eddy van Doorslaer. “Cut-Point Shift and Index Shift in Self-Reported Health.” Journal of Health Economics 23, no. 6 (November 1, 2004): 1083–99. https://doi.org/10.1016/j.jhealeco.2004.01.002.

MacLean, Paul S., Rena R. Wing, Terry Davidson, Leonard Epstein, Bret Goodpaster, Kevin D. Hall, Barry E. Levin, et al. “NIH Working Group Report: Innovative Research to Improve Maintenance of Weight Loss.” Obesity (Silver Spring, Md.) 23, no. 1 (January 2015): 7–15. https://doi.org/10.1002/oby.20967.

MacLean, Richard. “Organizational Design: Benchmarking.” Environmental Quality Management 22, no. 3 (2013): 95–108. https://doi.org/10.1002/tqem.21339.

McInerney, Melissa, Jennifer M. Mellor, and Lauren Hersch Nicholas. “Recession Depression: Mental Health Effects of the 2008 Stock Market Crash.” CESifo Working Paper Series. CESifo Working Paper Series. CESifo Group Munich, 2013. https://ideas.repec.org/p/ces/ceswps/_4263.html.

Miilunpalo, S., I. Vuori, P. Oja, M. Pasanen, and H. Urponen. “Self-Rated Health Status as a Health Measure: The Predictive Value of Self-Reported Health Status on the Use of Physician Services and on Mortality in the Working-Age Population.” Journal of Clinical Epidemiology 50, no. 5 (May 1997): 517–28. https://doi.org/10.1016/s0895-4356(97)00045-0.

Ruhm, Christopher, and William E. Black. “Does Drinking Really Decrease in Bad Times?” Journal of Health Economics 21, no. 4 (2002): 659–78.

Ruhm, Christopher J. “Healthy Living in Hard Times.” Journal of Health Economics 24, no. 2 (March 2005): 341–63. https://doi.org/10.1016/j.jhealeco.2004.09.007.

Vilares, Iris, James D. Howard, Hugo L. Fernandes, Jay A. Gottfried, and Konrad P. Kording. “Differential Representations of Prior and Likelihood Uncertainty in the Human Brain.” Current Biology: CB 22, no. 18 (September 25, 2012): 1641–48. https://doi.org/10.1016/j.cub.2012.07.010.

Vilares, Iris, and Konrad Kording. “Bayesian Models: The Structure of the World, Uncertainty, Behavior, and the Brain.” Annals of the New York Academy of Sciences 1224, no. 1 (April 2011): 22–39. https://doi.org/10.1111/j.1749-6632.2011.05965.x.


